### 2024-02-04 README.Rmd
add readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sfislands

<!-- badges: start -->
<!-- badges: end -->

The goal of sfislands is to ...

## Installation

You can install the development version of `kh` from [GitHub](https://github.com/) with:

```{r, echo=FALSE}

## Data and packages

# Read in data and necessary packages.

# install.packages("devtools")
devtools::install_github("horankev/sfislands")

# required packages
packages <- c(
  "tidyverse",
  "sf",
  "mgcv",
  "ggpubr",
  "spdep",
  "sfdep",
  "here"
)

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE, quietly = TRUE))

library(sfislands)

```

```{r, echo=FALSE}

### UK data

library(parlitools)

# prepare the data
# extract and join census and election data from parlitools package
census_11 <- parlitools::census_11 |> 
  select(-constituency_name,-constituency_type,-pano, -region, -country)
bes_2019 <- parlitools::bes_2019

elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))

# download constituency boundaries for 2019 using the st_read() command below
# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
# which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson

uk_map_download <- st_read(
  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
  quiet = TRUE)

# only need the boundaries and the IDs for merging with parlitools data
uk<- uk_map_download |> 
  select(pcon19cd,geometry) |> 
  st_transform(crs=27700) |> 
  st_make_valid() # ensure valid line overlaps etc

# create dataframe for filtering-out speaker constituencies below
speakers <- data.frame(
  year = c(2017,2019),
  constituency_name = c("Buckingham","Chorley")
)

# join constituency polygons to parlitools data
# filter out Scotland and Northern Ireland
# filter out the speakers' constituencies
# calculate the desired explanatory variables
# make levels into factors for use with `mgcv`

df <- elect_results |> 
  left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
  # filter(!country %in% c("Scotland","Northern Ireland")) |> 
  filter(!country %in% c("Northern Ireland")) |> 
  # filter(!constituency_name %in% speakers$constituency_name) |>
  mutate(degree_educated = qual_level_4,
         health_not_good = health_fair + health_bad + health_very_bad,
         white = ethnicity_white,
         con_change = con_19 - con_17, # difference in %
         lab_change = lab_19 - lab_17,
         con_swing = (con_change - lab_change)/2, # Butler swing
         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
         county = factor(county)) |> 
  st_as_sf()

# give the empty speaker seats the mean value, just for demostration purposes here
df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)

# create region boundaries layer for mapping
regions <- df |> group_by(region) |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# create county boundaries layer for mapping
counties <- df |> group_by(county) |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# create outline layer for mapping
uk_outline <- df |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# filter out required independent variables and scale them
# then add back to the features which were not scaled
df_scaled <- df |> 
  st_drop_geometry() |> 
  select(degree_educated,
         health_not_good,
         white) |> 
  scale() |> 
  as.data.frame() |> 
  mutate(con_swing = df$con_swing,
         population = df$population,
         region = factor(df$region),
         county = factor(df$county),
         constituency_name = df$constituency_name)

# create a simple features (spatial) version of this with a geometry column
df_scaled_sf <- df_scaled |> 
  mutate(geometry = df$geometry) |> 
  st_as_sf() |> 
  st_transform(crs=27700)
```

## Summary

1.  `sfislands` makes it easier to deal with geographic datasets which contain islands. These do not have to be "literal" islands but any situation where one or more discontiguous geographical unit is present. Such a situation can lead to two issues. Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible. Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent. The package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.

2.  As an aid to this, the package also has a function to quickly map any contiguity structure for visual inspection. This can also be used to examine the output of `sfdep` contiguity functions. This can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.

3.  If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.

4.  Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed. `sfdep` contains functionality to perform such test, and the output from `sfislands` can be used in its functions.

5.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.

6.  For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise. `sfislands` can streamline this workflow from the human side. Furthermore, there is a function to draw maps of these predictions for quick inspection.

## Functions overview

### 1. Set up data ("*pre-functions*")

-   **st_bridges()**

*Create a contiguity list, matrix, or `sf` dataframe with a contiguity list or matrix as column "nb", while accounting for islands.*

-   **st_quickmap_contigs()**

*Check contiguities visually on map.*

-   **st_check_islands()**

*Check assignment of island contiguities in a dataframe.*

-   **st_manual_join_nb()**

-   **st_manual_cut_nb()**

*Make manual changes to any connections.*

### 2. Create model

*Use the output of **st_bridges()** as both the data and contiguity inputs for a model using, for example, `mgcv`, `brms` or `inla`.*

### 3. Examine output ("*post functions*")

-   **st_augment()**

*Augment the original dataframe with model predictions.*

-   **st_quickmap_preds()**

*Generate quick maps of these predictions.*

## Pre-functions

`sfdep` offers excellent tools for building contiguity structures, among other things. It has a range of functions depending on how we want to define contiguity.

However, often when preparing areal spatial data, the presence of uncontiguous areas (such as islands or exclaves) can create difficulties. We might also want to account for some hidden contiguities by allowing bridges, tunnels etc. to render two uncontiguous areas as neighbours.

`sfislands` provides functions to make this task easier.

It also provides a number of further helper functions to examine these contiguity structures and use them in models such that the workflow is streamlined from the human side.

### Mainland France (`guerry` dataset)

`sfislands` is intended to be used alongside the `sfdep` package. The principal functions from `sfdep` for areal data are demonstrated in its vignette using the `guerry` dataset. This data applies to the geographical area of mainland France as shown below:

```{r}
g <- guerry |> 
  st_as_sf()

ggplot(g) + 
  geom_sf() + 
  theme_void()
```

In this French example, the `st_islands` function **st_bridges()** produces exactly the same contiguity structure as **sfdep::st_contiguity()**. This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_contigs()** which draws an easily editable map of the applied contiguity structure. This function works with contiguities constructed from any package as long as they are in list or matrix form.

For **st_quickmap_contigs()**, the contiguities should be within an `sf` dataframe as a column called "nb". **st_bridges()** does this automatically whereas the "nb" column needs to be added when using `sfdep` functions.

#### sfislands::st_bridges()

The `st_bridges()` contiguities below, where each department is considered a neighbour of another if it touches it at least one point...

```{r}
g |> st_bridges("department") |> 
  st_quickmap_contigs()

```

... are the same as these from `sfdep`:

#### sfdep::st_contiguity()

```{r}
g |> mutate(nb = st_contiguity(geometry)) |> 
  st_quickmap_contigs()

```

As these maps are produced using `ggplot2`, their characteristics can be edited using normal `ggplot2` syntax (see 1, below). For convenience, simple arguments are also provided (see 2, below) for changing the core characteristics of the map in a simple way:

```{r, fig.width=8}
ggarrange(

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_contigs() +
    geom_sf(data = g |> group_by(region) |> summarise(), 
            linewidth = 0.5, colour = "black", fill = NA) + 
    labs(title = "st_quickmap_contigs()",
         subtitle = "1. using ggplot syntax") + 
    theme_minimal() +
    theme(panel.background = element_rect(fill = "aquamarine3", color = "black"),
          axis.text = element_blank()),

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_contigs(linkcol = "orange", 
                        bordercol = "white", 
                        pointcol = "yellow", 
                        fillcol = "black", 
                        linksize = 0.4, 
                        bordersize = 0.3, 
                        pointsize = 0.8,
                        title = "st_quickmap_contigs()",
                        subtitle = "2. using simplified arguments"),

  ncol=2
)

```

`sfdep` offers a number of different types of contiguity structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_contigs()** function:

#### sfdep::st_dist_band()

All areas within a certain distance are considered neighbours:

```{r}

g |> mutate(nb = st_geometry(g) |> 
              st_dist_band(upper = 150000)) |> 
  st_quickmap_contigs()

```

#### sfdep::st_knn()

The k-nearest-neighbours (here, 1) to each area are considered neighbours:

```{r}

g |> mutate(nb = st_geometry(g) |> 
              st_knn(1, symmetric = TRUE)) |> 
  st_quickmap_contigs()

```

#### sfdep::st_block_nb()

All areas within a chosen block are considered neighbours:

```{r}

id <- g$code_dept
regime <- g$region
g |> 
  mutate(
    nb = st_block_nb(regime, id)
  ) |> 
  st_quickmap_contigs()

```

#### sfdep::st_lag_cumul()

Cumulative higher orders of contiguity such as also including neighbours-of-neighbours:

```{r}

g |> mutate(nb = st_contiguity(geometry) |> st_nb_lag_cumul(2)) |> 
  st_quickmap_contigs()

```

However, the `sfdep` functions above which rely on contiguity will run into difficulties if we consider the following geography:

### England, Scotland & Wales

In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in these graphs unless we attempt to pick out these islands and set up a buffer around them. This process is, however, cumbersome and can induce contiguities which are not intended.

One potential remedy is to entirely exclude islands from the study, another is to construct a contiguity structure according to your desired criteria and then set the islands to be contiguous to their closest *k* constituencies.

`st_bridges()` can do both of these things.

Below is the map in question:

```{r, echo=FALSE}
df_scaled_sf$country <- case_when(df_scaled_sf$region == "Wales" ~ "Wales",
                                  df_scaled_sf$region == "Scotland" ~ "Scotland",
                                  TRUE ~ "England")
```

There are island constituencies in the north around Scotland but also less obvious ones in Wales and England. The constituencies which are non-contiguous are outlined in red below:

```{r, fig.width=12, fig.height=10, echo=FALSE}
ggplot() + 
  geom_sf(data=df_scaled_sf, aes(fill=country), colour="white") + 
  # geom_sf_label(data=df_scaled_sf |> group_by(country) |> summarise(), aes(label=country),size=10) +
  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"),
                aes(label=constituency_name)) +
  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"),
                aes(label=constituency_name), nudge_x = -80000, nudge_y = -60000) +
    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),
                  aes(label=constituency_name), nudge_x = -90000) +
    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),
                  aes(label=constituency_name), nudge_x = 50000, nudge_y = -30000) +
  scale_fill_manual(values = c("antiquewhite4","#1C4769", "#24975E")) +
  # guides(fill = "none") + 
  theme_void()
```

To incorporate these we use `st_bridges()`. We can set `remove_islands` to TRUE if we decide to simply exclude the islands, or we set `link_k_islands` to the closest k constituencies to each islands which we want to bridge.

Below, with the argument `remove_islands` set to TRUE, we simply remove these islands from the dataset entirely.

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                  geom_col_name = "constituency_name",
                  remove_islands = T)
st_quickmap_contigs(nbsf,
                     pointsize=0.05,
                    title = "st_bridges() contiguities",
                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")

```

Alternatively, we can join islands to the nearest, say, 2 constituencies. `st_bridges()` by default returns the original dataframe augmented with a "nb" column which contains the contiguities in list form.

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                  geom_col_name = "constituency_name",
                  link_islands_k = 2)
st_quickmap_contigs(nbsf,
                     pointsize=0.05, 
                    title = "st_bridges() contiguities",
                      subtitle = "islands linked to k=2 nearest constituencies")

```

The contiguity structure which is created can be either a named list (the default) or a named matrix. Different modelling packages have different requirements for this. Furthermore, we can choose `add_to_dataframe` to be TRUE (the default) to return a dataframe with a column called `nb` which contains the named list or matrix. If FALSE, only the contiguity structure itself is returned

These options can be seen in the unexecuted code below:

```{r, eval=FALSE}
nbsf <- st_bridges(df = sf_dataframe,
                  geom_col_name = "the column containing the names of the contiguous areas",
                  remove_islands = T/F,
                  link_islands_k = 1...n,
                  nb_structure = "list"/"matrix",
                  add_to_dataframe = T/F,
                  title = "title",
                  subtitle = "subtitle")

```

The contiguity is here in a list form:

```{r}

head(nbsf$nb)

```

But the "nb" column can also be a matrix, and `st_quickmap_contigs()` will still return the same map:

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                  geom_col_name = "constituency_name",
                  remove_islands = F,
                  link_islands_k = 2,
                  nb_structure = "matrix",
                  add_to_dataframe = T)
st_quickmap_contigs(nbsf,
                     pointsize=0.05)

```

The matrix contiguity structure is now of the following form:

```{r}

nbsf$nb[1:10,1:10]

```

### Editing the contiguities

There are also functions for manually changing the results of a contiguity construction. It may be the case that you want to add some additional contiguities or to remove others. For example, you may be aware from local knowledge of connectivities which are not represented by mere contiguity of polygons. The presence of tunnels or bridges across a body of water would be an example of such a situation. The functions `st_manual_join_nb()` and `st_manual_cut_nb()` do this.

#### sfisland::st_check_islands()

To make the use of these functions easier and more intuitive, the function `st_check_islands()` shows us what contiguities have been set up for the islands by `st_bridges()`:

```{r}
nbsf |> st_check_islands()
```

Sometimes, as in this case, certain islands will have more than the specified k neighbours. This is due to the need for symmetry in the structure.

#### st_manual_join_nb() / st_manual_cut_nb()

Let us say we want to change some of these. For example, I will cut the tie between **Isle Of Wight** and **New Forest East** (using their numbers) and also between **Ynys Mon** and **Arfon** (using their names):

```{r}
nbsf |> 
  st_manual_cut_nb(292,378) |>
  st_manual_cut_nb("Ynys Mon","Arfon") |>
  st_check_islands()
```

As extra contiguities can be difficult to distinguish, an extreme case is shown below to demonstrate `st_manual_join_nb()`. **Gower** in South Wales is joined to **St Ives** in Cornwall and then mapped:

```{r}
st_bridges(df = df_scaled_sf|> filter(region %in% c("Wales","South West")),
           geom_col_name = "constituency_name",
           link_islands_k = 2
)  |> 
  st_manual_join_nb("Gower","St Ives") |>
  st_quickmap_contigs(title = "st_bridges contiguities: Wales & South West",
                      subtitle = "with additional st_manual_join_nb() for Gower and St Ives")
```

These manual functions can also, of course, be used to edit any of the previously discussed contiguity structures created by `sfdep`.

For example, looking just at Scotland we can use `sfdep::st_nb_lag_cumul()` to get first and second degree neighbours:

```{r}

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |>  
  st_quickmap_contigs()

```

This does not include the island constituencies. We can chose to include them by first using `st_bridges()` with k=2...

```{r}

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_quickmap_contigs()

```

... and then examining the connections which have been made for islands:

```{r}

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_check_islands()
```

We can then add these island connections to the output of `sfdep::st_nb_lag_cumul()`:

```{r}

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |> 
  st_manual_join_nb(47,9) |>
  st_manual_join_nb(47,51) |>
  st_manual_join_nb(47,55) |>
  st_manual_join_nb(51,9) |>
  st_manual_join_nb(51,47) |>
  st_quickmap_contigs()

```

## Modelling & post-functions

Having set up a contiguity structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.

For example, we can use the `mgcv` package to generate a Markov Random Field ICAR smooth of poor health across the study area. This is done very quickly by using `st_bridges()` to prepare the data, putting that inside the `mgcv` GAM formulation, and then piping into the `st_augment()` and `st_quickmap_preds()` functions.

```{r, fig.width=12, fig.height=10}

prep_data <- st_bridges(df_scaled_sf, "constituency_name")

gam(health_not_good ~ s(constituency_name, bs='mrf', xt=list(nb=prep_data$nb), k=100),
        data=prep_data, method="REML") |>
  st_augment(prep_data) |>
  st_quickmap_preds()

```

An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the contiguity to be in matrix form:

```{r, fig.width=12, fig.height=10}

prep_data2 <- st_bridges(df_scaled_sf, "constituency_name", nb_structure = "matrix")

# fit <- brm(degree_educated ~ car(W, gr=constituency_name, type="icar"),
#            data = prep_data2, data2 = list(W=prep_data2$nb),
#            family = gaussian(),file = "brmsfit_degree")

fit <- readRDS("brmsfit_degree.rds")

prep_data2$brmsfit <- predict(fit,prep_data2)[,1]

ggplot(prep_data2)+geom_sf(aes(fill=brmsfit), linewidth=0.1) +
  scale_fill_gradient2(low="firebrick4",mid="white",high="darkblue",midpoint = 0) +
  coord_sf(datum=NA) +
  theme_minimal() +
  theme_bw()

```

More complex models with random effects and multiple smooths are also possible with `mgcv` and the `st_augment()` and `st_quickmap_preds()` functions can handle these and label the columns and maps which are generated appropriately. This is shown below with a model of swing in the 2019 election:

```{r, fig.width=12, fig.height=6}

prep_data3 <- st_bridges(df_scaled_sf, "constituency_name") # decide upon the contiguities and add them to the df

model <- gam(con_swing ~ 
               s(region, bs="re") + # region level random intercept
               s(county, bs="re") + # county level random intercept
               s(county, degree_educated, bs="re") + # county level random coefficient
               s(constituency_name, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
               s(constituency_name, by=white, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
             data=prep_data3, method="REML") |> 
  st_augment(prep_data3) |> # pipe into function to get estimates
  st_quickmap_preds() # pipe into this for visualisation

ggarrange(plotlist = model, legend = "none", nrow=1)
```

To see the estimates returned:

```{r}

gam(con_swing ~ 
               s(region, bs="re") + # region level random intercept
               s(county, bs="re") + # county level random intercept
               s(county, degree_educated, bs="re") + # county level random coefficient
               s(constituency_name, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
               s(constituency_name, by=white, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
             data=prep_data3, method="REML") |> 
  st_augment(prep_data3) |> 
  head()

```

### Back to `guerry`

Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1831 as follows. Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a contiguity 'nb' column to the dataframe.

```{r}
prep_data4 <- g |> st_bridges("department")

mod4 <- gam(suicides ~ s(department, bs='mrf', xt=list(nb=prep_data4$nb), k=80),
            data=prep_data4, method="REML") |>
  st_augment(prep_data4) |>
  st_quickmap_preds()
ggarrange(plotlist=mod4)

```

Or fit a more complex mixed model:

```{r, fig.width=12, fig.height=6}
prep_data5 <- g |> 
  st_bridges("department")

model5 <- gam(donations ~ 
                s(region, bs="re") + 
                s(region, crime_prop, bs="re") + # county level random coefficient
                s(department, bs='mrf', 
                  xt=list(nb=prep_data5$nb),k=20) + # ICAR constituency level varying coefficents
                s(department, by=infants, bs='mrf', 
                  xt=list(nb=prep_data5$nb),k=20), # ICAR constituency level varying coefficents
              data=prep_data5, method="REML") |> 
  st_augment(prep_data5) |> # pipe into function to get estimates
  st_quickmap_preds() # pipe into this for visualisation

ggarrange(plotlist = model5, legend = "none", nrow = 1)
```

### 2024-02-04 README.Rmd
change description

@@
 
+``` r
+# install.packages("devtools")
+devtools::install_github("horankev/sfislands")
+```
+
 ```{r, echo=FALSE}
@@
 # install.packages("devtools")
-devtools::install_github("horankev/sfislands")
+# devtools::install_github("horankev/sfislands")
 
@@
 )  |> 
-  st_manual_join_nb("Gower","St Ives") |>
-  st_quickmap_contigs(title = "st_bridges contiguities: Wales & South West",
-                      subtitle = "with additional st_manual_join_nb() for Gower and St Ives")
+  st_manual_join_nb("Gower","St Ives")
 ```

### 2024-02-04 README.Rmd
change intro of readme

@@
 
-The goal of sfislands is to ...
+The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands.
+These do not have to be "literal" islands but any situation where one or more discontiguous geographical units are present.
+Such a situation can lead to two issues.
+Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.
+Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent.
+The package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.
 
@@
 ```
+
+## Summary of features
+
+1.
+As an aid to setting up contiguity structures, particularly when islands are involved, the package has a function to quickly map any contiguity structure for visual inspection.
+This can also be used to examine the output of `sfdep` contiguity functions.
+Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.
+
+2.
+If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
+
+3.
+Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed.
+`sfdep` contains functionality to perform such test, and the output from `sfislands` can be used in its functions.
+
+4.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.
+
+5.
+For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise.
+`sfislands` can streamline this workflow from the human side.
+Furthermore, there is a function to draw maps of these predictions for quick inspection.
+
+## Functions overview
+
+### 1. Set up data ("*pre-functions*")
+
+-   **st_bridges()**
+
+*Create a contiguity list, matrix, or `sf` dataframe with a contiguity list or matrix as column "nb", while accounting for islands.*
+
+-   **st_quickmap_contigs()**
+
+*Check contiguities visually on map.*
+
+-   **st_check_islands()**
+
+*Check assignment of island contiguities in a dataframe.*
+
+-   **st_manual_join_nb()**
+
+-   **st_manual_cut_nb()**
+
+*Make manual changes to any connections.*
+
+### 2. Create model
+
+*Use the output of **st_bridges()** as both the data and contiguity inputs for a model using, for example, `mgcv`, `brms` or `inla`.*
+
+### 3. Examine output ("*post functions*")
+
+-   **st_augment()**
+
+*Augment the original dataframe with model predictions.*
+
+-   **st_quickmap_preds()**
+
+*Generate quick maps of these predictions.*
 
@@
 ```
-
-## Summary
-
-1.
-`sfislands` makes it easier to deal with geographic datasets which contain islands.
-These do not have to be "literal" islands but any situation where one or more discontiguous geographical unit is present.
-Such a situation can lead to two issues.
-Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.
-Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent.
-The package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.
-
-2.
-As an aid to this, the package also has a function to quickly map any contiguity structure for visual inspection.
-This can also be used to examine the output of `sfdep` contiguity functions.
-This can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.
-
-3.
-If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
-
-4.
-Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed.
-`sfdep` contains functionality to perform such test, and the output from `sfislands` can be used in its functions.
-
-5.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.
-
-6.
-For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise.
-`sfislands` can streamline this workflow from the human side.
-Furthermore, there is a function to draw maps of these predictions for quick inspection.
-
-## Functions overview
-
-### 1. Set up data ("*pre-functions*")
-
--   **st_bridges()**
-
-*Create a contiguity list, matrix, or `sf` dataframe with a contiguity list or matrix as column "nb", while accounting for islands.*
-
--   **st_quickmap_contigs()**
-
-*Check contiguities visually on map.*
-
--   **st_check_islands()**
-
-*Check assignment of island contiguities in a dataframe.*
-
--   **st_manual_join_nb()**
-
--   **st_manual_cut_nb()**
-
-*Make manual changes to any connections.*
-
-### 2. Create model
-
-*Use the output of **st_bridges()** as both the data and contiguity inputs for a model using, for example, `mgcv`, `brms` or `inla`.*
-
-### 3. Examine output ("*post functions*")
-
--   **st_augment()**
-
-*Augment the original dataframe with model predictions.*
-
--   **st_quickmap_preds()**
-
-*Generate quick maps of these predictions.*
 

### 2024-02-04 README.Rmd
improve readme

@@
 The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands.
-These do not have to be "literal" islands but any situation where one or more discontiguous geographical units are present.
+These do not have to be "literal" islands but any situation where discontiguous geographical units are present.
 Such a situation can lead to two issues.
@@
 
-### 1. Set up data ("*pre-functions*")
+The following is a framework within which these functions could be used:
+
+### Step 1: Set up data ("*pre-functions*")
 
@@
 
-### 2. Create model
+### Step 2: Create model
 
@@
 
-### 3. Examine output ("*post functions*")
+### Step 3: Examine output ("*post functions*")
 
@@
 ```
+
+Below, we look at these functions in more detail and show them in operation.
 

### 2024-02-04 README.Rmd
improve readme

@@
 1.
+The initial setting up neighbourhood structures can be frustrating for people who are eager to get started with fitting spatial models.
+This is especially so when the presence of discontiguities within a geographical dataset means that, even having set up a neighbours list,  the model will still not run without further awkward data manipulations.
+
+2.
 As an aid to setting up contiguity structures, particularly when islands are involved, the package has a function to quickly map any contiguity structure for visual inspection.
@@
 
-2.
+3.
 If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
 
-3.
+4.
 Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed.
@@
 
-4.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.
+5.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.
 
-5.
+6.
 For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise.
@@
 
-The following is a framework within which these functions could be used:
+The following is a framework within which the `sfislands` functions could be used:
 

### 2024-02-04 README.Rmd
indents and bulleting of readme

@@
 <!-- badges: start -->
+
 <!-- badges: end -->
@@
 The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands.
-These do not have to be "literal" islands but any situation where discontiguous geographical units are present.
+
+-   These do not have to be "literal" islands but any situation where discontiguous geographical units are present.
+
 Such a situation can lead to two issues.
-Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.
-Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent.
-The package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.
+
+-   Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.
+
+-   Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent.
+
+This package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.
 

### 2024-02-05 README.Rmd
improvements to readme
add logo

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+
+# sfislands <img src="man/figures/logo.png" align="right" height="120" alt="" />
 
@@
 
--   Secondly, just because two areas are separated by, say, a body of water, this does not mean that they are necessarily independent.
+-   Secondly, just because two areas are separated by, say, a body of water, this does not necessarily mean that they are to be considered independent of each other.
 

### 2024-02-05 README.Rmd
improve readme

@@
 
-# sfislands <img src="man/figures/logo.png" align="right" height="120" alt="" />
+# sfislands <img src="man/figures/logo.png" align="right" height="120"/>
 
@@
 The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands.
+It does so using a tidy framework in the spirit of Josiah Parry's [sfdep](https://sfdep.josiahparry.com/) package.
 
@@
 `sfislands` is intended to be used alongside the `sfdep` package.
-The principal functions from `sfdep` for areal data are demonstrated in its vignette using the `guerry` dataset.
+The principal functions from `sfdep` for areal data are demonstrated in its vignette using the `guerry` ("*Essay on the Moral Statistics of France*") dataset.
 This data applies to the geographical area of mainland France as shown below:
@@
 
-Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1831 as follows.
+Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1830 as follows.
 Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a contiguity 'nb' column to the dataframe.

### 2024-02-05 README.Rmd
alter tabs

@@
            geom_col_name = "constituency_name",
-           link_islands_k = 2
-)  |> 
+           link_islands_k = 2)  |> 
   st_manual_join_nb("Gower","St Ives")

### 2024-02-05 R/st_quickmap_nb.R
change name quickmap_nb

@@ -15,7 +15,7 @@
 #' @export
 #'
 #' @examples
-st_quickmap_contigs <- function(nbsf,
+st_quickmap_nb <- function(nbsf,
                                 linkcol = "dodgerblue",
                                 bordercol = "gray7",
                                 pointcol="darkred",

### 2024-02-05 README.Rmd
nb not contig

@@
 2.
-As an aid to setting up contiguity structures, particularly when islands are involved, the package has a function to quickly map any contiguity structure for visual inspection.
-This can also be used to examine the output of `sfdep` contiguity functions.
+As an aid to setting up neighbourhood structures, particularly when islands are involved, the package has a function to quickly map any neighbourhood structure for visual inspection.
+This can also be used to examine the output of `sfdep` neighbour functions.
 Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.
 
-3.
-If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
+3.  If there are some neighbours assigned which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
 
@@
 
-*Create a contiguity list, matrix, or `sf` dataframe with a contiguity list or matrix as column "nb", while accounting for islands.*
+*Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.*
 
--   **st_quickmap_contigs()**
+-   **st_quickmap_nb()**
 
@@
 
-*Use the output of **st_bridges()** as both the data and contiguity inputs for a model using, for example, `mgcv`, `brms` or `inla`.*
+*Use the output of **st_bridges()** as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.*
 
@@
 
-`sfdep` offers excellent tools for building contiguity structures, among other things. It has a range of functions depending on how we want to define contiguity.
+`sfdep` offers excellent tools for building neighbourhood structures, among other things. It has a range of functions depending on how we want to define what *neighbour* should mean.
 
@@
 
-It also provides a number of further helper functions to examine these contiguity structures and use them in models such that the workflow is streamlined from the human side.
+It also provides a number of further helper functions to examine these neighbourhood structures and use them in models such that the workflow is streamlined from the human side.
 
@@
 
-In this French example, the `st_islands` function **st_bridges()** produces exactly the same contiguity structure as **sfdep::st_contiguity()**.
-This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_contigs()** which draws an easily editable map of the applied contiguity structure.
-This function works with contiguities constructed from any package as long as they are in list or matrix form.
+In this French example, the `st_islands` function **st_bridges()** produces exactly the same neighbourhood structure (based on contiguity) as **sfdep::st_contiguity()**.
+This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_nb()** which gives a visual representation of the structure on a map.
+This function works with neighbourhoods constructed from any package as long as they are in list or matrix form.
 
-For **st_quickmap_contigs()**, the contiguities should be within an `sf` dataframe as a column called "nb".
+For **st_quickmap_nb()**, the neighbourhoods should be within an `sf` dataframe as a column called "nb".
 **st_bridges()** does this automatically whereas the "nb" column needs to be added when using `sfdep` functions.
@@
 g |> st_bridges("department") |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
 g |> mutate(nb = st_contiguity(geometry)) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
   g |> mutate(nb = st_contiguity(geometry)) |> 
-    st_quickmap_contigs() +
+    st_quickmap_nb() +
     geom_sf(data = g |> group_by(region) |> summarise(), 
             linewidth = 0.5, colour = "black", fill = NA) + 
-    labs(title = "st_quickmap_contigs()",
+    labs(title = "st_quickmap_nb()",
          subtitle = "1. using ggplot syntax") + 
@@
   g |> mutate(nb = st_contiguity(geometry)) |> 
-    st_quickmap_contigs(linkcol = "orange", 
+    st_quickmap_nb(linkcol = "orange", 
                         bordercol = "white", 
@@
                         pointsize = 0.8,
-                        title = "st_quickmap_contigs()",
+                        title = "st_quickmap_nb()",
                         subtitle = "2. using simplified arguments"),
@@
 
-`sfdep` offers a number of different types of contiguity structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_contigs()** function:
+`sfdep` offers a number of different types of neighbourhood structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_nb()** function:
 
@@
               st_dist_band(upper = 150000)) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
               st_knn(1, symmetric = TRUE)) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
   ) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
 g |> mutate(nb = st_contiguity(geometry) |> st_nb_lag_cumul(2)) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
 
-In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in these graphs unless we attempt to pick out these islands and set up a buffer around them.
+In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in a contiguity graph unless we attempt to pick out these islands and set up a buffer around them.
 This process is, however, cumbersome and can induce contiguities which are not intended.
@@
                    remove_islands = T)
-st_quickmap_contigs(nbsf,
+st_quickmap_nb(nbsf,
                     pointsize=0.05,
@@
                    link_islands_k = 2)
-st_quickmap_contigs(nbsf,
+st_quickmap_nb(nbsf,
                     pointsize=0.05, 
@@
 
-The contiguity structure which is created can be either a named list (the default) or a named matrix.
-Different modelling packages have different requirements for this.
+The neighbourhood structure which is created can be either a named list (the default) or a named matrix.
+Different modelling packages have different requirements for how this information should be presented.
 Furthermore, we can choose `add_to_dataframe` to be TRUE (the default) to return a dataframe with a column called `nb` which contains the named list or matrix.
-If FALSE, only the contiguity structure itself is returned
+If FALSE, only the neighbourhood structure itself is returned
 
@@
 
-The contiguity is here in a list form:
+The neighbourhoods are here in a list form:
 
@@
 
-But the "nb" column can also be a matrix, and `st_quickmap_contigs()` will still return the same map:
+But the "nb" column can also be a matrix, and `st_quickmap_nb()` will still return the same map:
 
@@
                    add_to_dataframe = T)
-st_quickmap_contigs(nbsf,
+st_quickmap_nb(nbsf,
                     pointsize=0.05)
@@
 
-The matrix contiguity structure is now of the following form:
+The matrix neighbourhood structure is now of the following form:
 
@@
 
-There are also functions for manually changing the results of a contiguity construction.
-It may be the case that you want to add some additional contiguities or to remove others.
+There are also functions for manually changing the results of a neighbourhood construction.
+It may be the case that you want to add some additional links or to remove others.
 For example, you may be aware from local knowledge of connectivities which are not represented by mere contiguity of polygons.
@@
 
-These manual functions can also, of course, be used to edit any of the previously discussed contiguity structures created by `sfdep`.
+These manual functions can also, of course, be used to edit any of the previously discussed neighbourhood structures created by `sfdep`.
 
@@
            st_nb_lag_cumul(2)) |>  
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
              link_islands_k = 2) |> 
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
   st_manual_join_nb(51,47) |>
-  st_quickmap_contigs()
+  st_quickmap_nb()
 
@@
 
-Having set up a contiguity structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.
+Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.
 
@@
 
-An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the contiguity to be in matrix form:
+An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the neighbourhoods to be in matrix form:
 
@@
 Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1830 as follows.
-Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a contiguity 'nb' column to the dataframe.
+Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a neighbourhood 'nb' column to the dataframe.
 

### 2024-02-05 README.Rmd
remove refs to older name
remove duplication

@@
 
-# sfislands <img src="man/figures/logo.png" align="right" height="120"/>
+<img src="man/figures/logo.png" align="right" height="240"/>
 
@@
 
-You can install the development version of `kh` from [GitHub](https://github.com/) with:
+You can install the development version of `sfislands` from [GitHub](https://github.com/horankev/sfislands) with:
 

### 2024-02-05 README.Rmd
output as table

@@
 
--   **st_bridges()**
-
-*Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.*
-
--   **st_quickmap_nb()**
-
-*Check contiguities visually on map.*
-
--   **st_check_islands()**
-
-*Check assignment of island contiguities in a dataframe.*
-
--   **st_manual_join_nb()**
-
--   **st_manual_cut_nb()**
-
-*Make manual changes to any connections.*
+| function:               | purpose:                                                                                                                                   |
+|------------|-----------------------------------------------|
+| **st_bridges()**        | *Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.* |
+| **st_quickmap_nb()**    | *Check contiguities visually on map.*                                                                                                      |
+| **st_check_islands()**  | *Check assignment of island contiguities in a dataframe.*                                                                                  |
+| **st_manual_join_nb()** | *Make manual changes to any connections.*                                                                                                  |
+| **st_manual_cut_nb()**  | *Make manual changes to any connections.*                                                                                                  |
 
@@
 
--   **st_augment()**
-
-*Augment the original dataframe with model predictions.*
-
--   **st_quickmap_preds()**
-
-*Generate quick maps of these predictions.*
+| function:               | purpose:                                                 |
+|-----------------|-------------------------------------------------------|
+| **st_augment()**        | *Augment the original dataframe with model predictions.* |
+| **st_quickmap_preds()** | *Generate quick maps of these predictions.*              |
 

### 2024-02-05 R/st_bridges.R
function help

@@ -1,13 +1,13 @@
-#' Create contiguityies when islands are present
+#' Create contiguities when islands are present
 #'
-#' @param df
-#' @param geom_col_name
-#' @param remove_islands
-#' @param link_islands_k
-#' @param nb_structure
-#' @param add_to_dataframe
+#' @param df an sf or sfc object
+#' @param geom_col_name name of column containing names (or unique identifiers) for each row
+#' @param remove_islands TRUE or FALSE (default)
+#' @param link_islands_k an integer, k. Each island should be connected to the k nearest units
+#' @param nb_structure either "list" (default) or "matrix"
+#' @param add_to_dataframe TRUE (default) or FALSE. Whether or not to augment existing df with output as "nb" column
 #'
-#' @return
+#' @return a named neighbourhood list or matrix, optionally included as a column in an sf dataframe
 #' @export
 #'
 #' @examples
@@ -19,8 +19,8 @@ st_bridges <- function(df, # sf dataframe
                        add_to_dataframe = TRUE)
 {
 
-  if (!inherits(df,"sf")) {
-    stop("Error: This function requires a simple features dataframe as input")
+  if (!inherits(df,"sf") || !inherits(df,"sfc")) {
+    stop("Error: This function requires an sf or sfc object")
   }
 
   if (remove_islands == TRUE){

### 2024-02-05 R/st_augment.R
function argument descriptions

@@ -1,9 +1,9 @@
 #' Augment dataframe with predictions of model
 #'
-#' @param model
-#' @param df
+#' @param model an `mgcv` model
+#' @param df the sf data frame to be augmented with model preditions
 #'
-#' @return
+#' @return an augmented sd dataframe
 #' @export
 #'
 #' @examples

### 2024-02-05 R/st_check_islands.R

@@ -1,8 +1,8 @@
 #' Examine contiguity actions performed on islands
 #'
-#' @param data
+#' @param data an sf dataframe with a neighbourhood column called "nb"
 #'
-#' @return
+#' @return a dataframe reporting non-contiguous connections made by st_bridges()
 #' @export
 #'
 #' @examples

### 2024-02-05 R/st_manual_cut_nb.R

@@ -1,10 +1,10 @@
 #' Manual remove contiguity between two areas
 #'
-#' @param nb
-#' @param x
-#' @param y
+#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
+#' @param x name or number of first area
+#' @param y name or number of second area
 #'
-#' @return
+#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
 #' @export
 #'
 #' @examples

### 2024-02-05 R/st_manual_join_nb.R

@@ -1,10 +1,10 @@
 #' Manually enforce contiguity between two areas
 #'
-#' @param nb
-#' @param x
-#' @param y
+#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
+#' @param x name or number of first area
+#' @param y name or number of second area
 #'
-#' @return
+#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
 #' @export
 #'
 #' @examples

### 2024-02-05 R/st_quickmap_nb.R

@@ -1,17 +1,17 @@
 #' Visualise a contiguity structure on a map
 #'
-#' @param nbsf
-#' @param linkcol
-#' @param bordercol
-#' @param pointcol
-#' @param fillcol
-#' @param linksize
-#' @param bordersize
-#' @param pointsize
-#' @param title
-#' @param subtitle
+#' @param nbsf an sf dataframe with a neighbourhood column called "nb"
+#' @param linkcol colour of lines connecting neighbours
+#' @param bordercol colour of boundary lines between areas
+#' @param pointcol colour of centroid points
+#' @param fillcol fill of areas
+#' @param linksize linewidth of lines connecting neighbours
+#' @param bordersize linewidth of borders between areas
+#' @param pointsize sie of centroid points
+#' @param title plot title
+#' @param subtitle plot subtitle
 #'
-#' @return
+#' @return ggplot of areas and neighbourhood structure
 #' @export
 #'
 #' @examples

### 2024-02-05 R/st_quickmap_preds.R

@@ -1,8 +1,8 @@
 #' Visualilse the predictions of an `mgcv` model
 #'
-#' @param output
+#' @param output an augmented sf dataframe produced by st_augment()
 #'
-#' @return
+#' @return a list of ggplots
 #' @export
 #'
 #' @examples

### 2024-02-06 R/st_bridges.R
amend st_bridges warning

@@ -20,7 +20,7 @@ st_bridges <- function(df, # sf dataframe
 {
 
   if (!inherits(df,"sf") || !inherits(df,"sfc")) {
-    stop("Error: This function requires an sf or sfc object")
+    stop("Error: df must be an sf or sfc object")
   }
 
   if (remove_islands == TRUE){

### 2024-02-06 R/st_bridges.R
remove warning section again. not resolved

@@ -19,9 +19,9 @@ st_bridges <- function(df, # sf dataframe
                        add_to_dataframe = TRUE)
 {
 
-  if (!inherits(df,"sf") || !inherits(df,"sfc")) {
-    stop("Error: df must be an sf or sfc object")
-  }
+  # if (!inherits(df,"sf") || !inherits(df,"sfc")) {
+  #   stop("Error: df must be an sf or sfc object")
+  # }
 
   if (remove_islands == TRUE){
     # unconnected units

### 2024-02-06 README.Rmd
added subsections
with uk map data in inst/ext folder

@@
 
+# uk_map_download <- st_read(
+#   "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
+#   quiet = TRUE)
+
 uk_map_download <- st_read(
-  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
+  "inst/extdata/WPC_Dec_2019_UGCB_UK.shp",
   quiet = TRUE)
@@
 
+#### Remove islands
+
 Below, with the argument `remove_islands` set to TRUE, we simply remove these islands from the dataset entirely.
@@
 ```
+
+#### Connect islands
 
@@
 
-fit <- readRDS("brmsfit_degree.rds")
+fit <- readRDS("/Users/kevinhoran/Documents/sfislands_prepstuff/brmsfit_degree.rds")
 

### 2024-02-06 R/st_bridges.R
with data folder

@@ -11,6 +11,7 @@
 #' @export
 #'
 #' @examples
+#' uk_elections |> st_bridges("constituency_name")
 st_bridges <- function(df, # sf dataframe
                        geom_col_name,
                        remove_islands = FALSE,

### 2024-02-06 R/uk_election.R

@@ -1,0 +1,15 @@
+#' UK election data
+#'
+#' A subset of data from the World Health Organization Global Tuberculosis
+#' Report ...
+#'
+#' @format ## `uk_election`
+#' A data frame with 7,240 rows and 60 columns:
+#' \describe{
+#'   \item{country}{Country name}
+#'   \item{iso2, iso3}{2 & 3 letter ISO country codes}
+#'   \item{year}{Year}
+#'   ...
+#' }
+#' @source <https://www.who.int/teams/global-tuberculosis-programme/data>
+"uk_election"

### 2024-02-06 data-raw/uk_election.R

@@ -1,0 +1,78 @@
+## code to prepare `uk_election` dataset goes here
+
+library(parlitools)
+library(dplyr)
+library(sf)
+# prepare the data
+# extract and join census and election data from parlitools package
+census_11 <- parlitools::census_11 |>
+  dplyr::select(-constituency_name,-constituency_type,-pano, -region, -country)
+bes_2019 <- parlitools::bes_2019
+
+elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
+
+# download constituency boundaries for 2019 using the st_read() command below
+# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
+uk_map_download <- st_read(
+  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
+  quiet = TRUE)
+
+
+# only need the boundaries and the IDs for merging with parlitools data
+uk<- uk_map_download |>
+  dplyr::select(pcon19cd,geometry) |>
+  sf::st_transform(crs=27700) |>
+  sf::st_make_valid() # ensure valid line overlaps etc
+
+# create dataframe for filtering-out speaker constituencies below
+speakers <- data.frame(
+  year = c(2017,2019),
+  constituency_name = c("Buckingham","Chorley")
+)
+
+# join constituency polygons to parlitools data
+# filter out Scotland and Northern Ireland
+# filter out the speakers' constituencies
+# calculate the desired explanatory variables
+# make levels into factors for use with `mgcv`
+
+df <- elect_results |>
+  dplyr::left_join(uk, by=c("ons_const_id"="pcon19cd")) |>
+  # filter(!country %in% c("Scotland","Northern Ireland")) |>
+  dplyr::filter(!country %in% c("Northern Ireland")) |>
+  # filter(!constituency_name %in% speakers$constituency_name) |>
+  dplyr::mutate(degree_educated = qual_level_4,
+         health_not_good = health_fair + health_bad + health_very_bad,
+         white = ethnicity_white,
+         con_change = con_19 - con_17, # difference in %
+         lab_change = lab_19 - lab_17,
+         con_swing = (con_change - lab_change)/2, # Butler swing
+         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
+         county = factor(county)) |>
[…]

### 2024-02-06 R/st_augment.R
fix namespace, and add examples to function

@@ -7,6 +7,11 @@
 #' @export
 #'
 #' @examples
+#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands")) |>
+#' st_bridges("constituency_name")
+#' mgcv::gam(health_not_good ~
+#'   s(constituency_name, bs='mrf', xt=list(nb=ukdata$nb), k=100),data=ukdata, method="REML") |>
+#' st_augment(ukdata)
 st_augment <- function(model,df){
 
   if (!(class(model)[1] %in% "gam")) {

### 2024-02-06 R/st_check_islands.R

@@ -6,6 +6,9 @@
 #' @export
 #'
 #' @examples
+#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
+#' st_bridges(ukdata,"constituency_name") |>
+#' st_check_islands()
 st_check_islands <- function(data){
 
   if (is.data.frame(data) && !("nb" %in% colnames(data))) {

### 2024-02-06 R/st_manual_cut_nb.R

@@ -8,6 +8,10 @@
 #' @export
 #'
 #' @examples
+#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
+#' st_bridges(ukdata,"constituency_name") |>
+#' st_manual_cut_nb("Ynys Mon","Arfon") |>
+#' st_manual_cut_nb(292,378)
 st_manual_cut_nb <- function(nb,x,y){
 
   if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {

### 2024-02-06 R/st_manual_join_nb.R

@@ -8,6 +8,9 @@
 #' @export
 #'
 #' @examples
+#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
+#' st_bridges(ukdata,"constituency_name") |>
+#' st_manual_join_nb("Gower","St Ives")
 st_manual_join_nb <- function(nb,x,y){
 
   if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {

### 2024-02-06 R/st_quickmap_preds.R

@@ -6,6 +6,12 @@
 #' @export
 #'
 #' @examples
+#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands")) |>
+#' st_bridges("constituency_name")
+#' mgcv::gam(health_not_good ~
+#'   s(constituency_name, bs='mrf', xt=list(nb=ukdata$nb), k=100),data=ukdata, method="REML") |>
+#' st_augment(ukdata) |>
+#' st_quickmap_preds()
 st_quickmap_preds <- function(output){
 
   if (!inherits(output,"sf")) {

### 2024-02-06 R/uk_election.R
document the data

@@ -1,15 +1,24 @@
 #' UK election data
 #'
-#' A subset of data from the World Health Organization Global Tuberculosis
-#' Report ...
+#' Swing and socio-economic data for England, Scotland & Wales
+#' Census and voting data sourced from parlitools R package
+#' Spatial data sourced from UK government geoportal
 #'
 #' @format ## `uk_election`
-#' A data frame with 7,240 rows and 60 columns:
+#'    An sf and data.frame object with 632 rows and 9 columns
+#'
 #' \describe{
-#'   \item{country}{Country name}
-#'   \item{iso2, iso3}{2 & 3 letter ISO country codes}
-#'   \item{year}{Year}
+#'   \item{degree_educated}{percent of population with level 4 qualifications or higher}
+#'   \item{health_not_good}{percent of population reporting health to be fair, bad, or very bad}
+#'   \item{white}{percent of population of exclusively white ethnicity}
+#'   \item{con_swing}{Butler swing to the Conservative Party from the Labour Party from election 2019 to election 2019}
+#'   \item{population}{Constituency population}
+#'   \item{region}{Region}
+#'   \item{county}{County}
+#'   \item{constituency_name}{Constituency}
+#'   \item{geometry}{sfc polygons column}
 #'   ...
 #' }
-#' @source <https://www.who.int/teams/global-tuberculosis-programme/data>
+#'
+#' @source <https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk>, <https://docs.evanodell.com/parlitools/>
 "uk_election"

### 2024-02-06 README.Rmd
with data folder
remove inst/ext... and change function examples

@@
 
-### UK data
-
-library(parlitools)
-
-# prepare the data
-# extract and join census and election data from parlitools package
-census_11 <- parlitools::census_11 |> 
-  select(-constituency_name,-constituency_type,-pano, -region, -country)
-bes_2019 <- parlitools::bes_2019
-
-elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
-
-# download constituency boundaries for 2019 using the st_read() command below
-# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
-# which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson
-
+# ### UK data
+# 
+# library(parlitools)
+# 
+# # prepare the data
+# # extract and join census and election data from parlitools package
+# census_11 <- parlitools::census_11 |> 
+#   select(-constituency_name,-constituency_type,-pano, -region, -country)
+# bes_2019 <- parlitools::bes_2019
+# 
+# elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
+# 
+# # download constituency boundaries for 2019 using the st_read() command below
+# # they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
+# # which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson
+# 
 # uk_map_download <- st_read(
@@
 #   quiet = TRUE)
+# 
+# # only need the boundaries and the IDs for merging with parlitools data
+# uk<- uk_map_download |> 
+#   select(pcon19cd,geometry) |> 
+#   st_transform(crs=27700) |> 
+#   st_make_valid() # ensure valid line overlaps etc
+# 
+# # create dataframe for filtering-out speaker constituencies below
+# speakers <- data.frame(
+#   year = c(2017,2019),
+#   constituency_name = c("Buckingham","Chorley")
+# )
+# 
+# # join constituency polygons to parlitools data
+# # filter out Scotland and Northern Ireland
+# # filter out the speakers' constituencies
+# # calculate the desired explanatory variables
+# # make levels into factors for use with `mgcv`
+# 
+# df <- elect_results |> 
+#   left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
+#   # filter(!country %in% c("Scotland","Northern Ireland")) |> 
+#   filter(!country %in% c("Northern Ireland")) |> 
+#   # filter(!constituency_name %in% speakers$constituency_name) |>
+#   mutate(degree_educated = qual_level_4,
+#          health_not_good = health_fair + health_bad + health_very_bad,
+#          white = ethnicity_white,
+#          con_change = con_19 - con_17, # difference in %
+#          lab_change = lab_19 - lab_17,
+#          con_swing = (con_change - lab_change)/2, # Butler swing
+#          region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
+#          county = factor(county)) |> 
+#   st_as_sf()
+# 
+# # give the empty speaker seats the mean value, just for demostration purposes here
+# df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)
+# 
+# # create region boundaries layer for mapping
+# regions <- df |> group_by(region) |> 
+#   summarise() |> 
+#   st_as_sf() |> 
+#   st_transform(crs=27700)
+# 
+# # create county boundaries layer for mapping
+# counties <- df |> group_by(county) |> 
+#   summarise() |> 
+#   st_as_sf() |> 
+#   st_transform(crs=27700)
+# 
+# # create outline layer for mapping
+# uk_outline <- df |> 
+#   summarise() |> 
+#   st_as_sf() |> 
+#   st_transform(crs=27700)
+# 
+# # filter out required independent variables and scale them
+# # then add back to the features which were not scaled
+# df_scaled <- df |> 
+#   st_drop_geometry() |> 
+#   select(degree_educated,
+#          health_not_good,
+#          white) |> 
+#   scale() |> 
+#   as.data.frame() |> 
+#   mutate(con_swing = df$con_swing,
+#          population = df$population,
+#          region = factor(df$region),
+#          county = factor(df$county),
+#          constituency_name = df$constituency_name)
+# 
+# # create a simple features (spatial) version of this with a geometry column
+# df_scaled_sf <- df_scaled |> 
+#   mutate(geometry = df$geometry) |> 
+#   st_as_sf() |> 
+#   st_transform(crs=27700)
 
-uk_map_download <- st_read(
-  "inst/extdata/WPC_Dec_2019_UGCB_UK.shp",
-  quiet = TRUE)
-
-# only need the boundaries and the IDs for merging with parlitools data
-uk<- uk_map_download |> 
-  select(pcon19cd,geometry) |> 
-  st_transform(crs=27700) |> 
-  st_make_valid() # ensure valid line overlaps etc
-
-# create dataframe for filtering-out speaker constituencies below
-speakers <- data.frame(
-  year = c(2017,2019),
-  constituency_name = c("Buckingham","Chorley")
-)
-
-# join constituency polygons to parlitools data
-# filter out Scotland and Northern Ireland
-# filter out the speakers' constituencies
-# calculate the desired explanatory variables
-# make levels into factors for use with `mgcv`
-
-df <- elect_results |> 
-  left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
-  # filter(!country %in% c("Scotland","Northern Ireland")) |> 
-  filter(!country %in% c("Northern Ireland")) |> 
-  # filter(!constituency_name %in% speakers$constituency_name) |>
-  mutate(degree_educated = qual_level_4,
-         health_not_good = health_fair + health_bad + health_very_bad,
-         white = ethnicity_white,
-         con_change = con_19 - con_17, # difference in %
-         lab_change = lab_19 - lab_17,
-         con_swing = (con_change - lab_change)/2, # Butler swing
-         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
-         county = factor(county)) |> 
-  st_as_sf()
-
-# give the empty speaker seats the mean value, just for demostration purposes here
-df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)
-
-# create region boundaries layer for mapping
-regions <- df |> group_by(region) |> 
-  summarise() |> 
-  st_as_sf() |> 
-  st_transform(crs=27700)
-
-# create county boundaries layer for mapping
-counties <- df |> group_by(county) |> 
-  summarise() |> 
-  st_as_sf() |> 
-  st_transform(crs=27700)
-
-# create outline layer for mapping
-uk_outline <- df |> 
-  summarise() |> 
-  st_as_sf() |> 
-  st_transform(crs=27700)
-
-# filter out required independent variables and scale them
-# then add back to the features which were not scaled
-df_scaled <- df |> 
-  st_drop_geometry() |> 
-  select(degree_educated,
-         health_not_good,
-         white) |> 
-  scale() |> 
-  as.data.frame() |> 
-  mutate(con_swing = df$con_swing,
-         population = df$population,
-         region = factor(df$region),
-         county = factor(df$county),
-         constituency_name = df$constituency_name)
-
-# create a simple features (spatial) version of this with a geometry column
-df_scaled_sf <- df_scaled |> 
-  mutate(geometry = df$geometry) |> 
-  st_as_sf() |> 
-  st_transform(crs=27700)
+# df_scaled_sf <- uk_election
 ```
@@
 ```{r, echo=FALSE}
-df_scaled_sf$country <- case_when(df_scaled_sf$region == "Wales" ~ "Wales",
-                                  df_scaled_sf$region == "Scotland" ~ "Scotland",
+uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
+                                  uk_election$region == "Scotland" ~ "Scotland",
                                   TRUE ~ "England")
@@
 ggplot() + 
-  geom_sf(data=df_scaled_sf, aes(fill=country), colour="white") + 
-  # geom_sf_label(data=df_scaled_sf |> group_by(country) |> summarise(), aes(label=country),size=10) +
-  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
-  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"),
+  geom_sf(data=uk_election, aes(fill=country), colour="white") + 
+  # geom_sf_label(data=uk_election |> group_by(country) |> summarise(), aes(label=country),size=10) +
+  geom_sf(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
+  geom_sf_label(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"),
                 aes(label=constituency_name)) +
-  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
-  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"),
+  geom_sf(data=uk_election |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
+  geom_sf_label(data=uk_election |> filter(constituency_name == "Orkney and Shetland"),
                 aes(label=constituency_name), nudge_x = -80000, nudge_y = -60000) +
-    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
-    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),
+    geom_sf(data=uk_election |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
+    geom_sf_label(data=uk_election |> filter(constituency_name=="Ynys Mon"),
                   aes(label=constituency_name), nudge_x = -90000) +
-    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
-    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),
+    geom_sf(data=uk_election |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
+    geom_sf_label(data=uk_election |> filter(constituency_name=="Isle Of Wight"),
                   aes(label=constituency_name), nudge_x = 50000, nudge_y = -30000) +
@@
 ```{r, fig.width=12, fig.height=10}
-nbsf <- st_bridges(df = df_scaled_sf,
+nbsf <- st_bridges(df = uk_election,
                    geom_col_name = "constituency_name",
@@
 ```{r}
-st_bridges(df = df_scaled_sf|> filter(region %in% c("Wales","South West")),
+st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
            geom_col_name = "constituency_name",
@@
 
-df_scaled_sf |> 
+uk_election |> 
   filter(region == "Scotland") |> 
-  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
+  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
            st_nb_lag_cumul(2)) |>  
@@
 
-df_scaled_sf |> 
+uk_election |> 
   filter(region == "Scotland") |> 
@@
 
-df_scaled_sf |> 
+uk_election |> 
   filter(region == "Scotland") |> 
-  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
+  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
            st_nb_lag_cumul(2)) |> 
@@
 
-prep_data <- st_bridges(df_scaled_sf, "constituency_name")
+prep_data <- st_bridges(uk_election, "constituency_name")
 
@@
 
-prep_data2 <- st_bridges(df_scaled_sf, "constituency_name", nb_structure = "matrix")
+prep_data2 <- st_bridges(uk_election, "constituency_name", nb_structure = "matrix")
 
@@
 
-prep_data3 <- st_bridges(df_scaled_sf, "constituency_name") # decide upon the contiguities and add them to the df
+prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
 

### 2024-02-06 R/st_augment.R

@@ -7,11 +7,15 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands")) |>
-#' st_bridges("constituency_name")
+#' package_name <- "mgcv"
+#' if (!require(package_name, character.only = TRUE)) {
+#' install.packages(package_name)
+#' library(package_name, character.only = TRUE)
+#' }
+#' prepdata <- st_bridges(uk_election,"constituency_name")
 #' mgcv::gam(health_not_good ~
-#'   s(constituency_name, bs='mrf', xt=list(nb=ukdata$nb), k=100),data=ukdata, method="REML") |>
-#' st_augment(ukdata)
+#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
+#' st_augment(uk_election)
 st_augment <- function(model,df){
 
   if (!(class(model)[1] %in% "gam")) {

### 2024-02-06 R/st_bridges.R

@@ -11,8 +11,7 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
-#' st_bridges(ukdata,"constituency_name")
+#' st_bridges(uk_election,"constituency_name")
 st_bridges <- function(df, # sf dataframe
                        geom_col_name,
                        remove_islands = FALSE,

### 2024-02-06 R/st_check_islands.R

@@ -6,8 +6,7 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
-#' st_bridges(ukdata,"constituency_name") |>
+#' st_bridges(uk_election,"constituency_name") |>
 #' st_check_islands()
 st_check_islands <- function(data){
 

### 2024-02-06 R/st_manual_cut_nb.R

@@ -8,8 +8,7 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
-#' st_bridges(ukdata,"constituency_name") |>
+#' st_bridges(uk_election,"constituency_name") |>
 #' st_manual_cut_nb("Ynys Mon","Arfon") |>
 #' st_manual_cut_nb(292,378)
 st_manual_cut_nb <- function(nb,x,y){

### 2024-02-06 R/st_manual_join_nb.R

@@ -8,8 +8,7 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
-#' st_bridges(ukdata,"constituency_name") |>
+#' st_bridges(uk_election,"constituency_name") |>
 #' st_manual_join_nb("Gower","St Ives")
 st_manual_join_nb <- function(nb,x,y){
 

### 2024-02-06 R/st_quickmap_nb.R

@@ -15,8 +15,7 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands"))
-#' st_bridges(ukdata,"constituency_name") |>
+#' st_bridges(uk_election,"constituency_name") |>
 #' st_quickmap_nb()
 st_quickmap_nb <- function(nbsf,
                            linkcol = "dodgerblue",

### 2024-02-06 R/st_quickmap_preds.R

@@ -6,11 +6,10 @@
 #' @export
 #'
 #' @examples
-#' ukdata <- readRDS(system.file("extdata", "ukdata.rds", package="sfislands")) |>
-#' st_bridges("constituency_name")
+#' prepdata <- st_bridges(uk_election,"constituency_name")
 #' mgcv::gam(health_not_good ~
-#'   s(constituency_name, bs='mrf', xt=list(nb=ukdata$nb), k=100),data=ukdata, method="REML") |>
-#' st_augment(ukdata) |>
+#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
+#' st_augment(uk_election) |>
 #' st_quickmap_preds()
 st_quickmap_preds <- function(output){
 

### 2024-02-06 README.Rmd
remove data-building section from readme

@@
 
-```
-
-```{r, echo=FALSE}
-
-# ### UK data
-# 
-# library(parlitools)
-# 
-# # prepare the data
-# # extract and join census and election data from parlitools package
-# census_11 <- parlitools::census_11 |> 
-#   select(-constituency_name,-constituency_type,-pano, -region, -country)
-# bes_2019 <- parlitools::bes_2019
-# 
-# elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
-# 
-# # download constituency boundaries for 2019 using the st_read() command below
-# # they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
-# # which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson
-# 
-# uk_map_download <- st_read(
-#   "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
-#   quiet = TRUE)
-# 
-# # only need the boundaries and the IDs for merging with parlitools data
-# uk<- uk_map_download |> 
-#   select(pcon19cd,geometry) |> 
-#   st_transform(crs=27700) |> 
-#   st_make_valid() # ensure valid line overlaps etc
-# 
-# # create dataframe for filtering-out speaker constituencies below
-# speakers <- data.frame(
-#   year = c(2017,2019),
-#   constituency_name = c("Buckingham","Chorley")
-# )
-# 
-# # join constituency polygons to parlitools data
-# # filter out Scotland and Northern Ireland
-# # filter out the speakers' constituencies
-# # calculate the desired explanatory variables
-# # make levels into factors for use with `mgcv`
-# 
-# df <- elect_results |> 
-#   left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
-#   # filter(!country %in% c("Scotland","Northern Ireland")) |> 
-#   filter(!country %in% c("Northern Ireland")) |> 
-#   # filter(!constituency_name %in% speakers$constituency_name) |>
-#   mutate(degree_educated = qual_level_4,
-#          health_not_good = health_fair + health_bad + health_very_bad,
-#          white = ethnicity_white,
-#          con_change = con_19 - con_17, # difference in %
-#          lab_change = lab_19 - lab_17,
-#          con_swing = (con_change - lab_change)/2, # Butler swing
-#          region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
-#          county = factor(county)) |> 
-#   st_as_sf()
-# 
-# # give the empty speaker seats the mean value, just for demostration purposes here
-# df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)
-# 
-# # create region boundaries layer for mapping
-# regions <- df |> group_by(region) |> 
-#   summarise() |> 
-#   st_as_sf() |> 
-#   st_transform(crs=27700)
-# 
-# # create county boundaries layer for mapping
-# counties <- df |> group_by(county) |> 
-#   summarise() |> 
-#   st_as_sf() |> 
-#   st_transform(crs=27700)
-# 
-# # create outline layer for mapping
-# uk_outline <- df |> 
-#   summarise() |> 
-#   st_as_sf() |> 
-#   st_transform(crs=27700)
-# 
-# # filter out required independent variables and scale them
-# # then add back to the features which were not scaled
-# df_scaled <- df |> 
-#   st_drop_geometry() |> 
-#   select(degree_educated,
-#          health_not_good,
-#          white) |> 
-#   scale() |> 
-#   as.data.frame() |> 
-#   mutate(con_swing = df$con_swing,
-#          population = df$population,
-#          region = factor(df$region),
-#          county = factor(df$county),
-#          constituency_name = df$constituency_name)
-# 
-# # create a simple features (spatial) version of this with a geometry column
-# df_scaled_sf <- df_scaled |> 
-#   mutate(geometry = df$geometry) |> 
-#   st_as_sf() |> 
-#   st_transform(crs=27700)
-
-# df_scaled_sf <- uk_election
 ```
@@
            link_islands_k = 2)  |> 
-  st_manual_join_nb("Gower","St Ives")
+  st_manual_join_nb("Gower","St Ives") |> 
+  st_quickmap_nb()
 ```

### 2024-02-06 README.Rmd
shorten readme

@@
 
-Below, we look at these functions in more detail and show them in operation.
-
-## Pre-functions
-
-`sfdep` offers excellent tools for building neighbourhood structures, among other things. It has a range of functions depending on how we want to define what *neighbour* should mean.
-
-However, often when preparing areal spatial data, the presence of uncontiguous areas (such as islands or exclaves) can create difficulties.
-We might also want to account for some hidden contiguities by allowing bridges, tunnels etc.
-to render two uncontiguous areas as neighbours.
-
-`sfislands` provides functions to make this task easier.
-
-It also provides a number of further helper functions to examine these neighbourhood structures and use them in models such that the workflow is streamlined from the human side.
-
-### Mainland France (`guerry` dataset)
-
-`sfislands` is intended to be used alongside the `sfdep` package.
-The principal functions from `sfdep` for areal data are demonstrated in its vignette using the `guerry` ("*Essay on the Moral Statistics of France*") dataset.
-This data applies to the geographical area of mainland France as shown below:
-
-```{r}
-g <- guerry |> 
-  st_as_sf()
-
-ggplot(g) + 
-  geom_sf() + 
-  theme_void()
-```
-
-In this French example, the `st_islands` function **st_bridges()** produces exactly the same neighbourhood structure (based on contiguity) as **sfdep::st_contiguity()**.
-This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_nb()** which gives a visual representation of the structure on a map.
-This function works with neighbourhoods constructed from any package as long as they are in list or matrix form.
-
-For **st_quickmap_nb()**, the neighbourhoods should be within an `sf` dataframe as a column called "nb".
-**st_bridges()** does this automatically whereas the "nb" column needs to be added when using `sfdep` functions.
-
-#### sfislands::st_bridges()
-
-The `st_bridges()` contiguities below, where each department is considered a neighbour of another if it touches it at least one point...
-
-```{r}
-g |> st_bridges("department") |> 
-  st_quickmap_nb()
-
-```
-
-... are the same as these from `sfdep`:
-
-#### sfdep::st_contiguity()
-
-```{r}
-g |> mutate(nb = st_contiguity(geometry)) |> 
-  st_quickmap_nb()
-
-```
-
-As these maps are produced using `ggplot2`, their characteristics can be edited using normal `ggplot2` syntax (see 1, below).
-For convenience, simple arguments are also provided (see 2, below) for changing the core characteristics of the map in a simple way:
-
-```{r, fig.width=8}
-ggarrange(
-
-  g |> mutate(nb = st_contiguity(geometry)) |> 
-    st_quickmap_nb() +
-    geom_sf(data = g |> group_by(region) |> summarise(), 
-            linewidth = 0.5, colour = "black", fill = NA) + 
-    labs(title = "st_quickmap_nb()",
-         subtitle = "1. using ggplot syntax") + 
-    theme_minimal() +
-    theme(panel.background = element_rect(fill = "aquamarine3", color = "black"),
-          axis.text = element_blank()),
-
-  g |> mutate(nb = st_contiguity(geometry)) |> 
-    st_quickmap_nb(linkcol = "orange", 
-                        bordercol = "white", 
-                        pointcol = "yellow", 
-                        fillcol = "black", 
-                        linksize = 0.4, 
-                        bordersize = 0.3, 
-                        pointsize = 0.8,
-                        title = "st_quickmap_nb()",
-                        subtitle = "2. using simplified arguments"),
-
-  ncol=2
-)
-
-```
-
-`sfdep` offers a number of different types of neighbourhood structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_nb()** function:
-
-#### sfdep::st_dist_band()
-
-All areas within a certain distance are considered neighbours:
-
-```{r}
-
-g |> mutate(nb = st_geometry(g) |> 
-              st_dist_band(upper = 150000)) |> 
-  st_quickmap_nb()
-
-```
-
-#### sfdep::st_knn()
-
-The k-nearest-neighbours (here, 1) to each area are considered neighbours:
-
-```{r}
-
-g |> mutate(nb = st_geometry(g) |> 
-              st_knn(1, symmetric = TRUE)) |> 
-  st_quickmap_nb()
-
-```
-
-#### sfdep::st_block_nb()
-
-All areas within a chosen block are considered neighbours:
-
-```{r}
-
-id <- g$code_dept
-regime <- g$region
-g |> 
-  mutate(
-    nb = st_block_nb(regime, id)
-  ) |> 
-  st_quickmap_nb()
-
-```
-
-#### sfdep::st_lag_cumul()
-
-Cumulative higher orders of contiguity such as also including neighbours-of-neighbours:
-
-```{r}
-
-g |> mutate(nb = st_contiguity(geometry) |> st_nb_lag_cumul(2)) |> 
-  st_quickmap_nb()
-
-```
-
-However, the `sfdep` functions above which rely on contiguity will run into difficulties if we consider the following geography:
-
-### England, Scotland & Wales
-
-In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in a contiguity graph unless we attempt to pick out these islands and set up a buffer around them.
-This process is, however, cumbersome and can induce contiguities which are not intended.
-
-One potential remedy is to entirely exclude islands from the study, another is to construct a contiguity structure according to your desired criteria and then set the islands to be contiguous to their closest *k* constituencies.
-
-`st_bridges()` can do both of these things.
-
-Below is the map in question:
-
-```{r, echo=FALSE}
-uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
-                                  uk_election$region == "Scotland" ~ "Scotland",
-                                  TRUE ~ "England")
-```
-
-There are island constituencies in the north around Scotland but also less obvious ones in Wales and England. The constituencies which are non-contiguous are outlined in red below:
-
-```{r, fig.width=12, fig.height=10, echo=FALSE}
-ggplot() + 
-  geom_sf(data=uk_election, aes(fill=country), colour="white") + 
-  # geom_sf_label(data=uk_election |> group_by(country) |> summarise(), aes(label=country),size=10) +
-  geom_sf(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
-  geom_sf_label(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"),
-                aes(label=constituency_name)) +
-  geom_sf(data=uk_election |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
-  geom_sf_label(data=uk_election |> filter(constituency_name == "Orkney and Shetland"),
-                aes(label=constituency_name), nudge_x = -80000, nudge_y = -60000) +
-    geom_sf(data=uk_election |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
-    geom_sf_label(data=uk_election |> filter(constituency_name=="Ynys Mon"),
-                  aes(label=constituency_name), nudge_x = -90000) +
-    geom_sf(data=uk_election |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
-    geom_sf_label(data=uk_election |> filter(constituency_name=="Isle Of Wight"),
-                  aes(label=constituency_name), nudge_x = 50000, nudge_y = -30000) +
-  scale_fill_manual(values = c("antiquewhite4","#1C4769", "#24975E")) +
-  # guides(fill = "none") + 
-  theme_void()
-```
-
-To incorporate these we use `st_bridges()`.
-We can set `remove_islands` to TRUE if we decide to simply exclude the islands, or we set `link_k_islands` to the closest k constituencies to each islands which we want to bridge.
-
-#### Remove islands
-
-Below, with the argument `remove_islands` set to TRUE, we simply remove these islands from the dataset entirely.
-
-```{r, fig.width=12, fig.height=10}
-nbsf <- st_bridges(df = uk_election,
-                   geom_col_name = "constituency_name",
-                   remove_islands = T)
-st_quickmap_nb(nbsf,
-                    pointsize=0.05,
-                    title = "st_bridges() contiguities",
-                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")
-
-```
-
-#### Connect islands
-
-Alternatively, we can join islands to the nearest, say, 2 constituencies.
-`st_bridges()` by default returns the original dataframe augmented with a "nb" column which contains the contiguities in list form.
-
-```{r, fig.width=12, fig.height=10}
-nbsf <- st_bridges(df = uk_election,
-                   geom_col_name = "constituency_name",
-                   link_islands_k = 2)
-st_quickmap_nb(nbsf,
-                    pointsize=0.05, 
-                    title = "st_bridges() contiguities",
-                    subtitle = "islands linked to k=2 nearest constituencies")
-
-```
-
-The neighbourhood structure which is created can be either a named list (the default) or a named matrix.
-Different modelling packages have different requirements for how this information should be presented.
-Furthermore, we can choose `add_to_dataframe` to be TRUE (the default) to return a dataframe with a column called `nb` which contains the named list or matrix.
-If FALSE, only the neighbourhood structure itself is returned
-
-These options can be seen in the unexecuted code below:
-
-```{r, eval=FALSE}
-nbsf <- st_bridges(df = sf_dataframe,
-                  geom_col_name = "the column containing the names of the contiguous areas",
-                  remove_islands = T/F,
-                  link_islands_k = 1...n,
-                  nb_structure = "list"/"matrix",
-                  add_to_dataframe = T/F,
-                  title = "title",
-                  subtitle = "subtitle")
-
-```
-
-The neighbourhoods are here in a list form:
-
-```{r}
-
-head(nbsf$nb)
-
-```
-
-But the "nb" column can also be a matrix, and `st_quickmap_nb()` will still return the same map:
-
-```{r, fig.width=12, fig.height=10}
-nbsf <- st_bridges(df = uk_election,
-                   geom_col_name = "constituency_name",
-                   remove_islands = F,
-                   link_islands_k = 2,
-                   nb_structure = "matrix",
-                   add_to_dataframe = T)
-st_quickmap_nb(nbsf,
-                    pointsize=0.05)
-
-```
-
-The matrix neighbourhood structure is now of the following form:
-
-```{r}
-
-nbsf$nb[1:10,1:10]
-
-```
-
-### Editing the contiguities
-
-There are also functions for manually changing the results of a neighbourhood construction.
-It may be the case that you want to add some additional links or to remove others.
-For example, you may be aware from local knowledge of connectivities which are not represented by mere contiguity of polygons.
-The presence of tunnels or bridges across a body of water would be an example of such a situation.
-The functions `st_manual_join_nb()` and `st_manual_cut_nb()` do this.
-
-#### sfisland::st_check_islands()
-
-To make the use of these functions easier and more intuitive, the function `st_check_islands()` shows us what contiguities have been set up for the islands by `st_bridges()`:
-
-```{r}
-nbsf |> st_check_islands()
-```
-
-Sometimes, as in this case, certain islands will have more than the specified k neighbours. This is due to the need for symmetry in the structure.
-
-#### st_manual_join_nb() / st_manual_cut_nb()
-
-Let us say we want to change some of these.
-For example, I will cut the tie between **Isle Of Wight** and **New Forest East** (using their numbers) and also between **Ynys Mon** and **Arfon** (using their names):
-
-```{r}
-nbsf |> 
-  st_manual_cut_nb(292,378) |>
-  st_manual_cut_nb("Ynys Mon","Arfon") |>
-  st_check_islands()
-```
-
-As extra contiguities can be difficult to distinguish, an extreme case is shown below to demonstrate `st_manual_join_nb()`.
-**Gower** in South Wales is joined to **St Ives** in Cornwall and then mapped:
-
-```{r}
-st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
-           geom_col_name = "constituency_name",
-           link_islands_k = 2)  |> 
-  st_manual_join_nb("Gower","St Ives") |> 
-  st_quickmap_nb()
-```
-
-These manual functions can also, of course, be used to edit any of the previously discussed neighbourhood structures created by `sfdep`.
-
-For example, looking just at Scotland we can use `sfdep::st_nb_lag_cumul()` to get first and second degree neighbours:
-
-```{r}
-
-uk_election |> 
-  filter(region == "Scotland") |> 
-  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
-           st_nb_lag_cumul(2)) |>  
-  st_quickmap_nb()
-
-```
-
-This does not include the island constituencies. We can chose to include them by first using `st_bridges()` with k=2...
-
-```{r}
-
-uk_election |> 
-  filter(region == "Scotland") |> 
-  st_bridges(geom_col_name = "constituency_name",
-             link_islands_k = 2) |> 
-  st_quickmap_nb()
-
-```
-
-... and then examining the connections which have been made for islands:
-
-```{r}
-
-uk_election |> 
-  filter(region == "Scotland") |> 
-  st_bridges(geom_col_name = "constituency_name",
-             link_islands_k = 2) |> 
-  st_check_islands()
-```
-
-We can then add these island connections to the output of `sfdep::st_nb_lag_cumul()`:
-
-```{r}
-
-uk_election |> 
-  filter(region == "Scotland") |> 
-  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
-           st_nb_lag_cumul(2)) |> 
-  st_manual_join_nb(47,9) |>
-  st_manual_join_nb(47,51) |>
-  st_manual_join_nb(47,55) |>
-  st_manual_join_nb(51,9) |>
-  st_manual_join_nb(51,47) |>
-  st_quickmap_nb()
-
-```
-
-## Modelling & post-functions
-
-Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.
-
-For example, we can use the `mgcv` package to generate a Markov Random Field ICAR smooth of poor health across the study area.
-This is done very quickly by using `st_bridges()` to prepare the data, putting that inside the `mgcv` GAM formulation, and then piping into the `st_augment()` and `st_quickmap_preds()` functions.
-
-```{r, fig.width=12, fig.height=10}
-
-prep_data <- st_bridges(uk_election, "constituency_name")
-
-gam(health_not_good ~ s(constituency_name, bs='mrf', xt=list(nb=prep_data$nb), k=100),
-        data=prep_data, method="REML") |>
-  st_augment(prep_data) |>
-  st_quickmap_preds()
-
-```
-
-An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the neighbourhoods to be in matrix form:
-
-```{r, fig.width=12, fig.height=10}
-
-prep_data2 <- st_bridges(uk_election, "constituency_name", nb_structure = "matrix")
-
-# fit <- brm(degree_educated ~ car(W, gr=constituency_name, type="icar"),
-#            data = prep_data2, data2 = list(W=prep_data2$nb),
-#            family = gaussian(),file = "brmsfit_degree")
-
-fit <- readRDS("/Users/kevinhoran/Documents/sfislands_prepstuff/brmsfit_degree.rds")
-
-prep_data2$brmsfit <- predict(fit,prep_data2)[,1]
-
-ggplot(prep_data2)+geom_sf(aes(fill=brmsfit), linewidth=0.1) +
-  scale_fill_gradient2(low="firebrick4",mid="white",high="darkblue",midpoint = 0) +
-  coord_sf(datum=NA) +
-  theme_minimal() +
-  theme_bw()
-
-```
-
-More complex models with random effects and multiple smooths are also possible with `mgcv` and the `st_augment()` and `st_quickmap_preds()` functions can handle these and label the columns and maps which are generated appropriately.
-This is shown below with a model of swing in the 2019 election:
-
-```{r, fig.width=12, fig.height=6}
-
-prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
-
-model <- gam(con_swing ~ 
-               s(region, bs="re") + # region level random intercept
-               s(county, bs="re") + # county level random intercept
-               s(county, degree_educated, bs="re") + # county level random coefficient
-               s(constituency_name, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
-               s(constituency_name, by=white, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
-             data=prep_data3, method="REML") |> 
-  st_augment(prep_data3) |> # pipe into function to get estimates
-  st_quickmap_preds() # pipe into this for visualisation
-
-ggarrange(plotlist = model, legend = "none", nrow=1)
-```
-
-To see the estimates returned:
-
-```{r}
-
-gam(con_swing ~ 
-               s(region, bs="re") + # region level random intercept
-               s(county, bs="re") + # county level random intercept
-               s(county, degree_educated, bs="re") + # county level random coefficient
-               s(constituency_name, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
-               s(constituency_name, by=white, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
-             data=prep_data3, method="REML") |> 
-  st_augment(prep_data3) |> 
-  head()
-
-```
-
-### Back to `guerry`
-
-Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1830 as follows.
-Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a neighbourhood 'nb' column to the dataframe.
-
-```{r}
-prep_data4 <- g |> st_bridges("department")
-
-mod4 <- gam(suicides ~ s(department, bs='mrf', xt=list(nb=prep_data4$nb), k=80),
-            data=prep_data4, method="REML") |>
-  st_augment(prep_data4) |>
-  st_quickmap_preds()
-ggarrange(plotlist=mod4)
-
-```
-
-Or fit a more complex mixed model:
-
-```{r, fig.width=12, fig.height=6}
-prep_data5 <- g |> 
-  st_bridges("department")
-
-model5 <- gam(donations ~ 
-                s(region, bs="re") + 
-                s(region, crime_prop, bs="re") + # county level random coefficient
-                s(department, bs='mrf', 
-                  xt=list(nb=prep_data5$nb),k=20) + # ICAR constituency level varying coefficents
-                s(department, by=infants, bs='mrf', 
-                  xt=list(nb=prep_data5$nb),k=20), # ICAR constituency level varying coefficents
-              data=prep_data5, method="REML") |> 
-  st_augment(prep_data5) |> # pipe into function to get estimates
-  st_quickmap_preds() # pipe into this for visualisation
-
-ggarrange(plotlist = model5, legend = "none", nrow = 1)
-```
+More details of how to operate these functions are provided in the vignette.

### 2024-02-07 R/uk_election.R
change description of data file

@@ -8,14 +8,14 @@
 #'    An sf and data.frame object with 632 rows and 9 columns
 #'
 #' \describe{
-#'   \item{degree_educated}{percent of population with level 4 qualifications or higher}
-#'   \item{health_not_good}{percent of population reporting health to be fair, bad, or very bad}
-#'   \item{white}{percent of population of exclusively white ethnicity}
+#'   \item{degree_educated}{Percentage of constituency population with level 4 qualifications or higher, scaled to mean 0 and standard deviation 1}
+#'   \item{health_not_good}{Percentage of constituency of population reporting health to be fair, bad, or very bad, scaled to mean 0 and standard deviation 1}
+#'   \item{white}{Percentage of constituency of population of exclusively white ethnicity, scaled to mean 0 and standard deviation 1}
 #'   \item{con_swing}{Butler swing to the Conservative Party from the Labour Party from election 2019 to election 2019}
 #'   \item{population}{Constituency population}
-#'   \item{region}{Region}
-#'   \item{county}{County}
-#'   \item{constituency_name}{Constituency}
+#'   \item{region}{Regions}
+#'   \item{county}{Counties}
+#'   \item{constituency_name}{Westminster parliamentary constituencies, as of 2019}
 #'   \item{geometry}{sfc polygons column}
 #'   ...
 #' }

### 2024-02-07 README.Rmd
shorten readme, change themes, let logo be autoplaced

@@
 # sfislands
-
-<!-- badges: start -->
-
-<!-- badges: end -->
 
@@
 
-More details of how to operate these functions are provided in the vignette.
+## Examples
+
+In the context of the constituencies of England, Scotland and Wales, `st_bridges` will define a neighbourhood structure which includes the islands.
+`st_quickmap_nb` will show a representation of this on a map:
+
+```{r, echo=FALSE}
+uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
+                                  uk_election$region == "Scotland" ~ "Scotland",
+                                  TRUE ~ "England")
+```
+
+```{r, fig.width=12, fig.height=10}
+nbsf <- st_bridges(df = uk_election,
+                   geom_col_name = "constituency_name",
+                   link_islands_k = 2)
+st_quickmap_nb(nbsf,
+                    pointsize=0.05, 
+                    title = "st_bridges() contiguities",
+                    subtitle = "islands linked to k=2 nearest constituencies")
+
+```
+
+Neighbours can be manually added or removed using `st_manual_join_nb()` and `st_manual_cut_nb()`.
+
+```{r}
+st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
+           geom_col_name = "constituency_name",
+           link_islands_k = 2)  |> 
+  st_manual_join_nb("Gower","St Ives") |> 
+  st_quickmap_nb()
+```
+
+## Modelling & post-functions
+
+Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.
+
+For example, we can use the `mgcv` package to generate a Markov Random Field ICAR smooth of poor health across the study area.
+This is done very quickly by using `st_bridges()` to prepare the data, putting that inside the `mgcv` GAM formulation, and then piping into the `st_augment()` and `st_quickmap_preds()` functions.
+
+Further information and worked examples are available in this vignette.

### 2024-02-07 README.Rmd
remove duplicate logos
resize logo

@@
 
-<img src="man/figures/logo.png" align="right" height="240"/>
+<!-- <img src="man/figures/logo.png" align="right" height="240"/> -->
+<!-- # sfislands <a href="https://horankev.github.io/sfislands/"><img src="man/figures/logo.png" align="right" height="240" alt="sfislands website" /></a> -->
 
@@
 | function:               | purpose:                                                                                                                                   |
-|------------|-----------------------------------------------|
+|-----------------|-------------------------------------------------------|
 | **st_bridges()**        | *Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.* |
@@
 | function:               | purpose:                                                 |
-|-----------------|-------------------------------------------------------|
+|------------------|------------------------------------------------------|
 | **st_augment()**        | *Augment the original dataframe with model predictions.* |
@@
 
-Further information and worked examples are available in this vignette.
+Further information and worked examples are available in the vignettes which can be found [here](https://horankev.github.io/sfislands/).

### 2024-02-07 R/st_bridges.R
add orcid and change theme

@@ -1,4 +1,4 @@
-#' Create contiguities when islands are present
+#' Create neighbourhood structure when islands are present
 #'
 #' @param df an sf or sfc object
 #' @param geom_col_name name of column containing names (or unique identifiers) for each row

### 2024-02-07 R/st_quickmap_nb.R

@@ -1,4 +1,4 @@
-#' Visualise a contiguity structure on a map
+#' Visualise a neighbourhood structure on a map
 #'
 #' @param nbsf an sf dataframe with a neighbourhood column called "nb"
 #' @param linkcol colour of lines connecting neighbours

### 2024-02-07 README.Rmd
add a logo and change theme

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-
-<!-- <img src="man/figures/logo.png" align="right" height="240"/> -->
-<!-- # sfislands <a href="https://horankev.github.io/sfislands/"><img src="man/figures/logo.png" align="right" height="240" alt="sfislands website" /></a> -->
 
@@
 Furthermore, there is a function to draw maps of these predictions for quick inspection.
+
+<img src="man/figures/logo.png" align="right" height="240"/>
 
@@
 
+```{r, fig.width=12, fig.height=6}
+
+prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
+
+model <- gam(con_swing ~ 
+               s(region, bs="re") + # region level random intercept
+               s(county, bs="re") + # county level random intercept
+               s(county, degree_educated, bs="re") + # county level random coefficient
+               s(constituency_name, bs='mrf', 
+                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
+               s(constituency_name, by=white, bs='mrf', 
+                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
+             data=prep_data3, method="REML") |> 
+  st_augment(prep_data3) |> # pipe into function to get estimates
+  st_quickmap_preds() # pipe into this for visualisation
+
+ggarrange(plotlist = model, legend = "none", nrow=1)
+```
+
 Further information and worked examples are available in the vignettes which can be found [here](https://horankev.github.io/sfislands/).

### 2024-02-07 README.Rmd
move added logo

@@
 
-The following is a framework within which the `sfislands` functions could be used:
+The following is a framework within which the `sfislands` functions could be used.
+
+The first group can be seen as pre-functions, designed to create a neighbourhood structure suitable for certain types of models.
+
+Once a model has been fit (using `mgcv` in this case), the post-functions can be used to extract the predictions.
 

### 2024-02-07 README.Rmd
use saved png for output of brms model
correct sourcing of image from vignette/articles folder, labelling of joined constituencies

@@
   st_manual_join_nb("Gower","St Ives") |> 
-  st_quickmap_nb()
+  st_quickmap_nb() + 
+  geom_sf_label(data=uk_election|> filter(constituency_name %in% c("Gower","St Ives")),
+                aes(label=constituency_name),nudge_x = -30000)
 ```
@@
 
-Further information and worked examples are available in the vignettes which can be found [here](https://horankev.github.io/sfislands/).
+Further information and worked examples are available in the this [article](https://horankev.github.io/sfislands/).

### 2024-02-07 data-raw/uk_election.R
fix non-ascii warning

@@ -75,4 +75,7 @@ uk_election <- df_scaled |>
   sf::st_as_sf() |>
   sf::st_transform(crs=27700)
 
+# attempt to fix non-ascii character issue (degrees symbol)
+st_crs(uk_election)$wkt <- gsub("°|º", "\\\\u00b0", st_crs(uk_election)$wkt)
+
 usethis::use_data(uk_election, overwrite = TRUE, compress = "bzip2")

### 2024-02-18 R/st_quickmap_nb.R
numeric option for st_quickmap_nb

@@ -3,13 +3,16 @@
 #' @param nbsf an sf dataframe with a neighbourhood column called "nb"
 #' @param linkcol colour of lines connecting neighbours
 #' @param bordercol colour of boundary lines between areas
-#' @param pointcol colour of centroid points
+#' @param pointcol colour of centroid points if nodes == "point"
 #' @param fillcol fill of areas
 #' @param linksize linewidth of lines connecting neighbours
 #' @param bordersize linewidth of borders between areas
-#' @param pointsize sie of centroid points
+#' @param pointsize size of centroid points if nodes == "point"
 #' @param title plot title
 #' @param subtitle plot subtitle
+#' @param nodes point or numeric
+#' @param numericsize font size if nodes == "numeric"
+#' @param numericcol font colour if nodes == "numeric"
 #'
 #' @return ggplot of areas and neighbourhood structure
 #' @export
@@ -26,7 +29,10 @@ st_quickmap_nb <- function(nbsf,
                            bordersize=0.1,
                            pointsize=0.8,
                            title=NULL,
-                           subtitle=NULL){
+                           subtitle=NULL,
+                           nodes="point",
+                           numericsize=5,
+                           numericcol="black"){
 
   if (!inherits(nbsf,"sf")) {
     stop("Error: This function requires a simple features dataframe as input")
@@ -65,15 +71,35 @@ st_quickmap_nb <- function(nbsf,
   endpoints_coords <- sf::st_coordinates(neighbors_sf) |> data.frame() |>
     sf::st_as_sf(coords=c("X","Y"), crs=sf::st_crs(neighbors_sf))
 
-  # map the connections
-  ggplot2::ggplot() +
-    ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
-    ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-    ggplot2::geom_sf(data=endpoints_coords, size=pointsize, colour=pointcol) +
-    ggplot2::coord_sf(datum=NA) +
-    ggplot2::labs(title = title,
-                  subtitle = subtitle) +
-    ggplot2::theme_void() +
-    ggplot2::theme(axis.title.x = ggplot2::element_blank()) +
[…]

### 2024-02-18 R/st_quickmap_nb.R
colour and size for st_quickmap_nb numeric

@@ -93,8 +93,7 @@ st_quickmap_nb <- function(nbsf,
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
       ggplot2::geom_sf_text(data=endpoints_coords,
-                            aes(label=1:nrow(nbsf)), numericsize=numericsize, numericcol=numericcolour, fontface="bold") +
-      ggplot2::geom_sf(data=endpoints_coords, size=pointsize, colour=pointcol) +
+                            ggplot2::aes(label=rownames(endpoints_coords)), numericsize=numericsize, numericcol=numericcol, fontface="bold") +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fix numeric issues

@@ -3,16 +3,16 @@
 #' @param nbsf an sf dataframe with a neighbourhood column called "nb"
 #' @param linkcol colour of lines connecting neighbours
 #' @param bordercol colour of boundary lines between areas
-#' @param pointcol colour of centroid points if nodes == "point"
+#' @param pointcol colour of centroid points if nodes are point
 #' @param fillcol fill of areas
 #' @param linksize linewidth of lines connecting neighbours
 #' @param bordersize linewidth of borders between areas
-#' @param pointsize size of centroid points if nodes == "point"
+#' @param pointsize size of centroid points if nodes are point
 #' @param title plot title
 #' @param subtitle plot subtitle
 #' @param nodes point or numeric
-#' @param numericsize font size if nodes == "numeric"
-#' @param numericcol font colour if nodes == "numeric"
+#' @param numericsize font size if nodes are numeric
+#' @param numericcol font colour if nodes are numeric
 #'
 #' @return ggplot of areas and neighbourhood structure
 #' @export
@@ -71,7 +71,22 @@ st_quickmap_nb <- function(nbsf,
   endpoints_coords <- sf::st_coordinates(neighbors_sf) |> data.frame() |>
     sf::st_as_sf(coords=c("X","Y"), crs=sf::st_crs(neighbors_sf))
 
-  if(nodes == "point"){
+  if(nodes == "numeric"){
+
+    # map the connections
+    ggplot2::ggplot() +
+      ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
+      ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
+      ggplot2::geom_sf_text(data=endpoints_coords, label=1:nrow(endpoints_coords), numericsize=numericsize, numericcol=numericcol, fontface="bold") +
+      ggplot2::coord_sf(datum=NA) +
+      ggplot2::labs(title = title,
+                    subtitle = subtitle) +
+      ggplot2::theme_void() +
+      ggplot2::theme(axis.title.x = ggplot2::element_blank()) +
+      ggplot2::theme(axis.title.y = ggplot2::element_blank())
+  }
+
+  else{
 
     # map the connections
     ggplot2::ggplot() +
@@ -86,19 +101,4 @@ st_quickmap_nb <- function(nbsf,
[…]

### 2024-02-18 R/st_quickmap_nb.R
try to fix numeric bit

@@ -31,7 +31,7 @@ st_quickmap_nb <- function(nbsf,
                            title=NULL,
                            subtitle=NULL,
                            nodes="point",
-                           numericsize=5,
+                           numericsize=1,
                            numericcol="black"){
 
   if (!inherits(nbsf,"sf")) {
@@ -77,7 +77,7 @@ st_quickmap_nb <- function(nbsf,
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, label=1:nrow(endpoints_coords), numericsize=numericsize, numericcol=numericcol, fontface="bold") +
+      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=1:nrow(endpoints_coords)), numericsize=numericsize, numericcol=numericcol, fontface="bold") +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fixes

@@ -73,11 +73,13 @@ st_quickmap_nb <- function(nbsf,
 
   if(nodes == "numeric"){
 
+    endpoints_coords$id <- 1:nrow(endpoints_coords)
+
     # map the connections
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=1:nrow(endpoints_coords)), numericsize=numericsize, numericcol=numericcol, fontface="bold") +
+      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=!!id), numericsize=numericsize, numericcol=numericcol) +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fixes

@@ -79,7 +79,7 @@ st_quickmap_nb <- function(nbsf,
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=!!id), numericsize=numericsize, numericcol=numericcol) +
+      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=endpoints_coords$id), numericsize=numericsize, numericcol=numericcol) +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fixes

@@ -79,7 +79,7 @@ st_quickmap_nb <- function(nbsf,
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=endpoints_coords$id), numericsize=numericsize, numericcol=numericcol) +
+      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=endpoints_coords$id), size=numericsize, colour=numericcol) +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fixes

@@ -79,7 +79,7 @@ st_quickmap_nb <- function(nbsf,
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=endpoints_coords$id), size=numericsize, colour=numericcol) +
+      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=id), size=numericsize, colour=numericcol) +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fix

@@ -73,13 +73,13 @@ st_quickmap_nb <- function(nbsf,
 
   if(nodes == "numeric"){
 
-    endpoints_coords$id <- 1:nrow(endpoints_coords)
+    nbsf$id <- 1:nrow(nbsf)
 
     # map the connections
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=endpoints_coords, aes(label=id), size=numericsize, colour=numericcol) +
+      ggplot2::geom_sf_text(data=nbsf |> st_centroid(), aes(label=id), size=numericsize, colour=numericcol) +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-18 R/st_quickmap_nb.R
fix

@@ -79,7 +79,7 @@ st_quickmap_nb <- function(nbsf,
     ggplot2::ggplot() +
       ggplot2::geom_sf(data=nbsf, fill=fillcol, colour=bordercol, linewidth=bordersize) +
       ggplot2::geom_sf(data = neighbors_sf, colour=linkcol, linewidth=linksize) +
-      ggplot2::geom_sf_text(data=nbsf |> st_centroid(), aes(label=id), size=numericsize, colour=numericcol) +
+      ggplot2::geom_sf_text(data=nbsf |> st_centroid(), aes(label=id), size=numericsize, colour=numericcol, fontface="bold") +
       ggplot2::coord_sf(datum=NA) +
       ggplot2::labs(title = title,
                     subtitle = subtitle) +

### 2024-02-22 R/st_quickmap_preds.R
add to quickmap preds

@@ -5,6 +5,9 @@
 #' @param scale_mid fill of midpoint of scale
 #' @param scale_high fill of highest extreme of scale
 #' @param scale_midpoint value of midpoint of scale
+#' @param borderwidth linewidth of borders between units
+#' @param bordercol colour of borders between units
+#' @param limits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
 #'
 #' @return a list of ggplots
 #' @export
@@ -19,7 +22,10 @@ st_quickmap_preds <- function(output,
                               scale_low = "firebrick4",
                               scale_mid = "white",
                               scale_high = "darkblue",
-                              scale_midpoint = 0){
+                              scale_midpoint = 0,
+                              borderwidth = 0.05,
+                              bordercol = "black",
+                              legendlimits = "individual"){
 
   if (!inherits(output,"sf")) {
     stop("Error: This function requires a simple features dataframe as input")
@@ -27,6 +33,14 @@ st_quickmap_preds <- function(output,
 
   output1 <- output |>
     dplyr::select(dplyr::starts_with("random.effect"),dplyr::starts_with("mrf.smooth"))
+
+  min_scale <- output1 |>
+    sf::st_drop_geometry() |>
+    min(na.rm = TRUE)
+
+  max_scale <- output1 |>
+    sf::st_drop_geometry() |>
+    max(na.rm = TRUE)
 
   fillnames <- output1 |>
     sf::st_drop_geometry() |>
@@ -41,11 +55,14 @@ st_quickmap_preds <- function(output,
   plot_list <- list()
   for (i in 1:length(fillnames)){
     plot_list[[i]] <- ggplot2::ggplot() +
-      ggplot2::geom_sf(data=output1, ggplot2::aes(fill=!!as.name(fillnames[i])), linewidth=0.05, colour="black") +
+      ggplot2::geom_sf(data=output1, ggplot2::aes(fill=!!as.name(fillnames[i])), linewidth=borderwidth, colour=bordercol) +
       ggplot2::scale_fill_gradient2(low = scale_low,
                                     mid = scale_mid,
                                     high = scale_high,
-                                    midpoint = scale_midpoint) +
+                                    midpoint = scale_midpoint,
+                                    limits = ifelse(legendlimits == "minmax",
+                                                    c(min_scale, max_scale),
+                                                    NULL)) +
       ggplot2::labs(title=newtitle[i],
[…]

### 2024-02-22 R/st_quickmap_preds.R
add args to quickmap preds

@@ -7,7 +7,7 @@
 #' @param scale_midpoint value of midpoint of scale
 #' @param borderwidth linewidth of borders between units
 #' @param bordercol colour of borders between units
-#' @param limits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
+#' @param legendlimits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
 #'
 #' @return a list of ggplots
 #' @export
@@ -60,9 +60,8 @@ st_quickmap_preds <- function(output,
                                     mid = scale_mid,
                                     high = scale_high,
                                     midpoint = scale_midpoint,
-                                    limits = ifelse(legendlimits == "minmax",
-                                                    c(min_scale, max_scale),
-                                                    NULL)) +
+                                    limits = if(legendlimits == "minmax") c(min_scale, max_scale)
+                                                else(NULL)) +
       ggplot2::labs(title=newtitle[i],
                     subtitle=newsubtitle[i]) +
       ggplot2::coord_sf(datum=NA) +

### 2024-03-03 R/st_augment.R
some tests

@@ -1,83 +1,0 @@
-#' Augment dataframe with predictions of model
-#'
-#' @param model an `mgcv` model
-#' @param df the sf data frame to be augmented with model preditions
-#'
-#' @return an augmented sd dataframe
-#' @export
-#'
-#' @examples
-#' package_name <- "mgcv"
-#' if (!require(package_name, character.only = TRUE)) {
-#' install.packages(package_name)
-#' library(package_name, character.only = TRUE)
-#' }
-#' prepdata <- st_bridges(uk_election,"constituency_name")
-#' mgcv::gam(health_not_good ~
-#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
-#' st_augment(uk_election)
-st_augment <- function(model,df){
-
-  if (!(class(model)[1] %in% "gam")) {
-    stop("Error: This function requires the model to be of class 'gam'")
-  }
-
-  # Identify numeric columns
-  numeric_columns <- sapply(df, is.numeric)
-
-  # Replace numeric columns with the number 1
-  tempdf <- df
-  tempdf[, numeric_columns] <- 1
-
-  output <- stats::predict(model, tempdf, type = "terms", se.fit = TRUE) |>
-    as.data.frame()
-
-  ### need different process for renaming cols if one versus more than one smooth:
-
-  ### if only one smooth:
-
-  if(ncol(output) == 2 & nrow(summary(model$smooth)) == 1) {
-    # change names from fit. to the type of effect (random effect or mrf.smooth)
-    names(output)[1] <- paste0(summary(model$smooth)[,2],".")
-    names(output)[stringr::str_starts(names(output),"fit.s.")] <- stringr::str_replace(names(output)[stringr::str_starts(names(output),"fit.s.")],
-                                                                                       "fit.s.",
-                                                                                       paste0(summary(model$smooth)[,2],"."))
-    # same for standard error columns
-    names(output)[2] <- paste0("se.",summary(model$smooth)[,2],".")
-  }
-
-  ### if more than one smooth:
-
-  else {
-
-    # change names from fit. to the type of effect (random effect or mrf.smooth)
-    names(output)[stringr::str_starts(names(output),"fit.s.")] <- stringr::str_replace(names(output)[stringr::str_starts(names(output),"fit.s.")],
-                                                                                       "fit.s.",
[…]

### 2024-03-03 R/st_bridges.R

@@ -1,113 +1,0 @@
-#' Create neighbourhood structure when islands are present
-#'
-#' @param df an sf or sfc object
-#' @param geom_col_name name of column containing names (or unique identifiers) for each row
-#' @param remove_islands TRUE or FALSE (default)
-#' @param link_islands_k an integer, k. Each island should be connected to the k nearest units
-#' @param nb_structure either "list" (default) or "matrix"
-#' @param add_to_dataframe TRUE (default) or FALSE. Whether or not to augment existing df with output as "nb" column
-#'
-#' @return a named neighbourhood list or matrix, optionally included as a column in an sf dataframe
-#' @export
-#'
-#' @examples
-#' st_bridges(uk_election,"constituency_name")
-st_bridges <- function(df, # sf dataframe
-                       geom_col_name,
-                       remove_islands = FALSE,
-                       link_islands_k = 1, # link island to k nearest units, 0 removes all islands
-                       nb_structure = "list",
-                       add_to_dataframe = TRUE)
-{
-
-  # if (!inherits(df,"sf") || !inherits(df,"sfc")) {
-  #   stop("Error: df must be an sf or sfc object")
-  # }
-
-  if (remove_islands == TRUE){
-    # unconnected units
-    cont <-  df |>
-      sf::st_intersects()
-    still_unconnected <- lengths(lapply(cont, function(x) x))
-    unconnected <- which(still_unconnected == 1)
-
-    df <- df[-unconnected,]
-    link_islands_k = 0
-
-  }
-
-  if (link_islands_k > 0)
-  {
-    # unconnected units
-    cont <-  df |>
-      sf::st_intersects()
-    still_unconnected <- lengths(lapply(cont, function(x) x))
-    unconnected <- which(still_unconnected == 1)
-    # Calculate distances
-    distdf <- data.frame(
-      constnumb = 0,
-      ndist = 0
-    )
-    for (i in 1:length(unconnected))
-    {
-      distances <- sf::st_distance(df$geometry[unconnected[i]], sf::st_geometry(df$geometry)) |>
-        as.numeric() |> sort()
-      distdf[i,2] <- distances[link_islands_k+1] * 1.001
-      distdf[i,1] <- unconnected[i]
-    }
[…]

### 2024-03-03 R/st_check_islands.R

@@ -1,50 +1,0 @@
-#' Examine contiguity actions performed on islands
-#'
-#' @param data an sf dataframe with a neighbourhood column called "nb"
-#'
-#' @return a dataframe reporting non-contiguous connections made by st_bridges()
-#' @export
-#'
-#' @examples
-#' st_bridges(uk_election,"constituency_name") |>
-#' st_check_islands()
-st_check_islands <- function(data){
-
-  if (is.data.frame(data) && !("nb" %in% colnames(data))) {
-    stop("Error: The dataframe must contain a column called 'nb'")
-  }
-
-  if (!(is.list(data$nb) || is.matrix(data$nb))) {
-    stop("Error: The 'nb' argument must be a neighbours list or a neighbours matrix")
-  }
-
-  # make it a list if it's a matrix
-  if(is.matrix(data$nb)){ # matrix within dataframe
-    keep_rownames <- rownames(data$nb)
-    tempnb <- spdep::mat2listw(data$nb, style="B")
-    tempnb <- tempnb[2]
-    data$nb <- tempnb$neighbours
-    names(data$nb) <- keep_rownames
-  }
-
-  # unconnected units
-  cont <-  data |>
-    sf::st_intersects()
-  still_unconnected <- lengths(lapply(cont, function(x) x))
-  unconnected <- which(still_unconnected == 1)
-
-
-  output <- data[unconnected,]$nb
-  names(output) <- paste(names(output)," : ",unconnected)
-
-  output2 <- do.call(rbind, lapply(names(output), function(x) cbind(x, output[[x]]))) |>
-    data.frame()
-  names(output2) <- c("region","nb_num")
-
-  consnames <- names(data$nb)
-  output2$nb_names <- consnames[as.integer(output2$nb_num)]
-
-  output2 <- tidyr::separate(output2, "region", into = c("island_names", "island_num"), sep = " : ")
-
-  return(output2)
-}

### 2024-03-03 R/st_manual_cut_nb.R

@@ -1,138 +1,0 @@
-#' Manual remove contiguity between two areas
-#'
-#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
-#' @param x name or number of first area
-#' @param y name or number of second area
-#'
-#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
-#' @export
-#'
-#' @examples
-#' st_bridges(uk_election,"constituency_name") |>
-#' st_manual_cut_nb("Ynys Mon","Arfon") |>
-#' st_manual_cut_nb(292,378)
-st_manual_cut_nb <- function(nb,x,y){
-
-  if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
-    stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
-  } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
-    stop("Error: The dataframe must contain a column called 'nb'")
-  }
-
-  ### case when numeric x and y:
-  # xnum<-x |> as.integer()
-  # ynum<-y |> as.integer()
-  if(is.numeric(x)){xnum <- x |> as.integer()}
-  if(is.numeric(y)){ynum <- y |> as.integer()}
-
-  ### case when character x and y:
-  # first, make into numeric is given as character
-  # depending on its structure as df, list, matrix...
-  if(is.character(x) & is.character(y)){
-
-    if(is.data.frame(nb)){
-      if(is.character(x)){
-        if(is.list(nb$nb)){
-          xnum <- which(names(nb$nb)==x) |>
-            as.integer()
-        }
-        if(is.matrix(nb$nb)){
-          xnum <- which(rownames(nb$nb)==x) |>
-            as.integer()
-        }
-      }
-      if(is.character(y)){
-        if(is.list(nb$nb)){
-          ynum <- which(names(nb$nb)==y) |>
-            as.integer()
-        }
-        if(is.matrix(nb$nb)){
-          ynum <- which(rownames(nb$nb)==y) |>
-            as.integer()
-        }
-      }
-    }else{
-      if(is.list(nb)){
-        xnum <- which(names(nb)==x) |>
-          as.integer()
[…]

### 2024-03-03 R/st_manual_join_nb.R

@@ -1,136 +1,0 @@
-#' Manually enforce contiguity between two areas
-#'
-#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
-#' @param x name or number of first area
-#' @param y name or number of second area
-#'
-#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
-#' @export
-#'
-#' @examples
-#' st_bridges(uk_election,"constituency_name") |>
-#' st_manual_join_nb("Gower","St Ives")
-st_manual_join_nb <- function(nb,x,y){
-
-  if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
-    stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
-  } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
-    stop("Error: The dataframe must contain a column called 'nb'")
-  }
-
-  # MAKE INTEGERS: xnum and ynum
-
-  ### case when numeric x and y:
-  if(is.numeric(x) & is.numeric(y)){
-    xnum <- x |> as.integer()
-    ynum <- y |> as.integer()
-  }else{
-
-    ### case when character x and y:
-    # first, make into numeric is given as character
-    # depending on its structure as df, list, matrix...
-    if(is.data.frame(nb)){
-      if(is.list(nb$nb)){
-        xnum <- which(names(nb$nb)==x) |>
-          as.integer()
-        ynum <- which(names(nb$nb)==y) |>
-          as.integer()
-      }
-      if(is.matrix(nb$nb)){
-        xnum <- which(rownames(nb$nb)==x) |>
-          as.integer()
-        ynum <- which(rownames(nb$nb)==y) |>
-          as.integer()
-      }
-    }else{
-      if(is.list(nb)){
-        xnum <- which(names(nb)==x) |>
-          as.integer()
-        ynum <- which(names(nb)==y) |>
-          as.integer()
-      }
-      if(is.matrix(nb)){
-        xnum <- which(rownames(nb)==x) |>
-          as.integer()
-        ynum <- which(rownames(nb)==y) |>
-          as.integer()
-      }
-    }
[…]

### 2024-03-03 R/st_quickmap_nb.R

@@ -1,151 +1,0 @@
-#' Visualise a neighbourhood structure on a map
-#'
-#' @param nbsf an sf dataframe with a neighbourhood column called "nb"
-#' @param linkcol colour of lines connecting neighbours
-#' @param bordercol colour of boundary lines between areas
-#' @param pointcol colour of centroid points if nodes are point
-#' @param fillcol fill of areas
-#' @param linksize linewidth of lines connecting neighbours
-#' @param bordersize linewidth of borders between areas
-#' @param pointsize size of centroid points if nodes are point
-#' @param title plot title
-#' @param subtitle plot subtitle
-#' @param nodes point or numeric
-#' @param numericsize font size if nodes are numeric
-#' @param numericcol font colour if nodes are numeric
-#' @param concavehull default FALSE. whether or not to show concave hulls
-#' @param hullratio fraction convex: 1 returns the convex hulls, 0 maximally concave hulls
-#' @param hullcol colour of concave hull lines
-#' @param hullsize line width of concave hull lines
-#'
-#' @return ggplot of areas and neighbourhood structure
-#' @export
-#'
-#' @examples
-#' st_bridges(uk_election,"constituency_name") |>
-#' st_quickmap_nb()
-st_quickmap_nb <- function(nbsf,
-                           linkcol = "dodgerblue",
-                           bordercol = "gray7",
-                           pointcol="darkred",
-                           fillcol = "gray95",
-                           linksize=0.2,
-                           bordersize=0.1,
-                           pointsize=0.8,
-                           title=NULL,
-                           subtitle=NULL,
-                           nodes="point",
-                           numericsize=5,
-                           numericcol="black",
-                           concavehull=FALSE,
-                           hullratio=0.8,
-                           hullcol="darkgreen",
-                           hullsize=0.5){
-
-  if (!inherits(nbsf,"sf")) {
-    stop("Error: This function requires a simple features dataframe as input")
-  }
-
[…]

### 2024-03-03 R/st_quickmap_preds.R

@@ -1,92 +1,0 @@
-#' Visualilse the predictions of an `mgcv` model
-#'
-#' @param output an augmented sf dataframe produced by st_augment()
-#' @param scale_low fill of lowest extreme of scale
-#' @param scale_mid fill of midpoint of scale
-#' @param scale_high fill of highest extreme of scale
-#' @param scale_midpoint value of midpoint of scale
-#' @param borderwidth linewidth of borders between units
-#' @param bordercol colour of borders between units
-#' @param legendlimits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
-#' @param titlesize font size for title
-#' @param subtitlesize font size for subtitle
-#' @param framefill colour for background fill
-#' @param frameline colour for frame
-#' @param framesize line width of frame
-#'
-#' @return a list of ggplots
-#' @export
-#'
-#' @examples
-#' prepdata <- st_bridges(uk_election,"constituency_name")
-#' mgcv::gam(health_not_good ~
-#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
-#' st_augment(uk_election) |>
-#' st_quickmap_preds()
-st_quickmap_preds <- function(output,
-                              scale_low = "firebrick4",
-                              scale_mid = "white",
-                              scale_high = "darkblue",
-                              scale_midpoint = 0,
-                              borderwidth = 0.05,
-                              bordercol = "black",
-                              legendlimits = "individual",
-                              titlesize = 12,
-                              subtitlesize = 10,
-                              framefill = "white",
-                              frameline = "black",
-                              framesize = 1){
-
-  if (!inherits(output,"sf")) {
-    stop("Error: This function requires a simple features dataframe as input")
-  }
-
-  output1 <- output |>
-    dplyr::select(dplyr::starts_with("random.effect"),dplyr::starts_with("mrf.smooth"))
-
-  min_scale <- output1 |>
-    sf::st_drop_geometry() |>
-    min(na.rm = TRUE)
-
-  max_scale <- output1 |>
-    sf::st_drop_geometry() |>
-    max(na.rm = TRUE)
-
-  fillnames <- output1 |>
-    sf::st_drop_geometry() |>
-    names()
-
-  # split column names into title and subtitle
-  # either side of second . in string
-  newsubtitle <- sub("^(.*?\\..*?)\\..*$", "\\1", fillnames)
[…]

### 2024-03-03 R/uk_election.R

@@ -1,24 +1,0 @@
-#' UK election data
-#'
-#' Swing and socio-economic data for England, Scotland & Wales
-#' Census and voting data sourced from parlitools R package
-#' Spatial data sourced from UK government geoportal
-#'
-#' @format ## `uk_election`
-#'    An sf and data.frame object with 632 rows and 9 columns
-#'
-#' \describe{
-#'   \item{degree_educated}{Percentage of constituency population with level 4 qualifications or higher, scaled to mean 0 and standard deviation 1}
-#'   \item{health_not_good}{Percentage of constituency of population reporting health to be fair, bad, or very bad, scaled to mean 0 and standard deviation 1}
-#'   \item{white}{Percentage of constituency of population of exclusively white ethnicity, scaled to mean 0 and standard deviation 1}
-#'   \item{con_swing}{Butler swing to the Conservative Party from the Labour Party from election 2019 to election 2019}
-#'   \item{population}{Constituency population}
-#'   \item{region}{Regions}
-#'   \item{county}{Counties}
-#'   \item{constituency_name}{Westminster parliamentary constituencies, as of 2019}
-#'   \item{geometry}{sfc polygons column}
-#'   ...
-#' }
-#'
-#' @source <https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk>, <https://docs.evanodell.com/parlitools/>
-"uk_election"

### 2024-03-03 data-raw/uk_election.R

@@ -1,81 +1,0 @@
-## code to prepare `uk_election` dataset goes here
-
-library(parlitools)
-library(dplyr)
-library(sf)
-# prepare the data
-# extract and join census and election data from parlitools package
-census_11 <- parlitools::census_11 |>
-  dplyr::select(-constituency_name,-constituency_type,-pano, -region, -country)
-bes_2019 <- parlitools::bes_2019
-
-elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
-
-# download constituency boundaries for 2019 using the st_read() command below
-# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
-uk_map_download <- st_read(
-  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
-  quiet = TRUE)
-
-
-# only need the boundaries and the IDs for merging with parlitools data
-uk<- uk_map_download |>
-  dplyr::select(pcon19cd,geometry) |>
-  sf::st_transform(crs=27700) |>
-  sf::st_make_valid() # ensure valid line overlaps etc
-
-# create dataframe for filtering-out speaker constituencies below
-speakers <- data.frame(
-  year = c(2017,2019),
-  constituency_name = c("Buckingham","Chorley")
-)
-
-# join constituency polygons to parlitools data
-# filter out Scotland and Northern Ireland
-# filter out the speakers' constituencies
-# calculate the desired explanatory variables
-# make levels into factors for use with `mgcv`
-
-df <- elect_results |>
-  dplyr::left_join(uk, by=c("ons_const_id"="pcon19cd")) |>
-  # filter(!country %in% c("Scotland","Northern Ireland")) |>
-  dplyr::filter(!country %in% c("Northern Ireland")) |>
-  # filter(!constituency_name %in% speakers$constituency_name) |>
-  dplyr::mutate(degree_educated = qual_level_4,
-         health_not_good = health_fair + health_bad + health_very_bad,
-         white = ethnicity_white,
-         con_change = con_19 - con_17, # difference in %
-         lab_change = lab_19 - lab_17,
-         con_swing = (con_change - lab_change)/2, # Butler swing
-         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
[…]

### 2024-03-03 R/st_augment.R
this?

@@ -1,0 +1,83 @@
+#' Augment dataframe with predictions of model
+#'
+#' @param model an `mgcv` model
+#' @param df the sf data frame to be augmented with model preditions
+#'
+#' @return an augmented sd dataframe
+#' @export
+#'
+#' @examples
+#' package_name <- "mgcv"
+#' if (!require(package_name, character.only = TRUE)) {
+#' install.packages(package_name)
+#' library(package_name, character.only = TRUE)
+#' }
+#' prepdata <- st_bridges(uk_election,"constituency_name")
+#' mgcv::gam(health_not_good ~
+#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
+#' st_augment(uk_election)
+st_augment <- function(model,df){
+
+  if (!(class(model)[1] %in% "gam")) {
+    stop("Error: This function requires the model to be of class 'gam'")
+  }
+
+  # Identify numeric columns
+  numeric_columns <- sapply(df, is.numeric)
+
+  # Replace numeric columns with the number 1
+  tempdf <- df
+  tempdf[, numeric_columns] <- 1
+
+  output <- stats::predict(model, tempdf, type = "terms", se.fit = TRUE) |>
+    as.data.frame()
+
+  ### need different process for renaming cols if one versus more than one smooth:
+
+  ### if only one smooth:
+
+  if(ncol(output) == 2 & nrow(summary(model$smooth)) == 1) {
+    # change names from fit. to the type of effect (random effect or mrf.smooth)
+    names(output)[1] <- paste0(summary(model$smooth)[,2],".")
+    names(output)[stringr::str_starts(names(output),"fit.s.")] <- stringr::str_replace(names(output)[stringr::str_starts(names(output),"fit.s.")],
+                                                                                       "fit.s.",
+                                                                                       paste0(summary(model$smooth)[,2],"."))
+    # same for standard error columns
+    names(output)[2] <- paste0("se.",summary(model$smooth)[,2],".")
+  }
+
+  ### if more than one smooth:
+
+  else {
+
+    # change names from fit. to the type of effect (random effect or mrf.smooth)
+    names(output)[stringr::str_starts(names(output),"fit.s.")] <- stringr::str_replace(names(output)[stringr::str_starts(names(output),"fit.s.")],
+                                                                                       "fit.s.",
[…]

### 2024-03-03 R/st_bridges.R

@@ -1,0 +1,113 @@
+#' Create neighbourhood structure when islands are present
+#'
+#' @param df an sf or sfc object
+#' @param geom_col_name name of column containing names (or unique identifiers) for each row
+#' @param remove_islands TRUE or FALSE (default)
+#' @param link_islands_k an integer, k. Each island should be connected to the k nearest units
+#' @param nb_structure either "list" (default) or "matrix"
+#' @param add_to_dataframe TRUE (default) or FALSE. Whether or not to augment existing df with output as "nb" column
+#'
+#' @return a named neighbourhood list or matrix, optionally included as a column in an sf dataframe
+#' @export
+#'
+#' @examples
+#' st_bridges(uk_election,"constituency_name")
+st_bridges <- function(df, # sf dataframe
+                       geom_col_name,
+                       remove_islands = FALSE,
+                       link_islands_k = 1, # link island to k nearest units, 0 removes all islands
+                       nb_structure = "list",
+                       add_to_dataframe = TRUE)
+{
+
+  # if (!inherits(df,"sf") || !inherits(df,"sfc")) {
+  #   stop("Error: df must be an sf or sfc object")
+  # }
+
+  if (remove_islands == TRUE){
+    # unconnected units
+    cont <-  df |>
+      sf::st_intersects()
+    still_unconnected <- lengths(lapply(cont, function(x) x))
+    unconnected <- which(still_unconnected == 1)
+
+    df <- df[-unconnected,]
+    link_islands_k = 0
+
+  }
+
+  if (link_islands_k > 0)
+  {
+    # unconnected units
+    cont <-  df |>
+      sf::st_intersects()
+    still_unconnected <- lengths(lapply(cont, function(x) x))
+    unconnected <- which(still_unconnected == 1)
+    # Calculate distances
+    distdf <- data.frame(
+      constnumb = 0,
+      ndist = 0
+    )
+    for (i in 1:length(unconnected))
+    {
+      distances <- sf::st_distance(df$geometry[unconnected[i]], sf::st_geometry(df$geometry)) |>
+        as.numeric() |> sort()
+      distdf[i,2] <- distances[link_islands_k+1] * 1.001
+      distdf[i,1] <- unconnected[i]
+    }
[…]

### 2024-03-03 R/st_check_islands.R

@@ -1,0 +1,50 @@
+#' Examine contiguity actions performed on islands
+#'
+#' @param data an sf dataframe with a neighbourhood column called "nb"
+#'
+#' @return a dataframe reporting non-contiguous connections made by st_bridges()
+#' @export
+#'
+#' @examples
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_check_islands()
+st_check_islands <- function(data){
+
+  if (is.data.frame(data) && !("nb" %in% colnames(data))) {
+    stop("Error: The dataframe must contain a column called 'nb'")
+  }
+
+  if (!(is.list(data$nb) || is.matrix(data$nb))) {
+    stop("Error: The 'nb' argument must be a neighbours list or a neighbours matrix")
+  }
+
+  # make it a list if it's a matrix
+  if(is.matrix(data$nb)){ # matrix within dataframe
+    keep_rownames <- rownames(data$nb)
+    tempnb <- spdep::mat2listw(data$nb, style="B")
+    tempnb <- tempnb[2]
+    data$nb <- tempnb$neighbours
+    names(data$nb) <- keep_rownames
+  }
+
+  # unconnected units
+  cont <-  data |>
+    sf::st_intersects()
+  still_unconnected <- lengths(lapply(cont, function(x) x))
+  unconnected <- which(still_unconnected == 1)
+
+
+  output <- data[unconnected,]$nb
+  names(output) <- paste(names(output)," : ",unconnected)
+
+  output2 <- do.call(rbind, lapply(names(output), function(x) cbind(x, output[[x]]))) |>
+    data.frame()
+  names(output2) <- c("region","nb_num")
+
+  consnames <- names(data$nb)
+  output2$nb_names <- consnames[as.integer(output2$nb_num)]
+
+  output2 <- tidyr::separate(output2, "region", into = c("island_names", "island_num"), sep = "  : ")
+
+  return(output2)
+}

### 2024-03-03 R/st_manual_cut_nb.R

@@ -1,0 +1,138 @@
+#' Manual remove contiguity between two areas
+#'
+#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
+#' @param x name or number of first area
+#' @param y name or number of second area
+#'
+#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
+#' @export
+#'
+#' @examples
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_manual_cut_nb("Ynys Mon","Arfon") |>
+#' st_manual_cut_nb(292,378)
+st_manual_cut_nb <- function(nb,x,y){
+
+  if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
+    stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
+  } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
+    stop("Error: The dataframe must contain a column called 'nb'")
+  }
+
+  ### case when numeric x and y:
+  # xnum<-x |> as.integer()
+  # ynum<-y |> as.integer()
+  if(is.numeric(x)){xnum <- x |> as.integer()}
+  if(is.numeric(y)){ynum <- y |> as.integer()}
+
+  ### case when character x and y:
+  # first, make into numeric is given as character
+  # depending on its structure as df, list, matrix...
+  if(is.character(x) & is.character(y)){
+
+    if(is.data.frame(nb)){
+      if(is.character(x)){
+        if(is.list(nb$nb)){
+          xnum <- which(names(nb$nb)==x) |>
+            as.integer()
+        }
+        if(is.matrix(nb$nb)){
+          xnum <- which(rownames(nb$nb)==x) |>
+            as.integer()
+        }
+      }
+      if(is.character(y)){
+        if(is.list(nb$nb)){
+          ynum <- which(names(nb$nb)==y) |>
+            as.integer()
+        }
+        if(is.matrix(nb$nb)){
+          ynum <- which(rownames(nb$nb)==y) |>
+            as.integer()
+        }
+      }
+    }else{
+      if(is.list(nb)){
+        xnum <- which(names(nb)==x) |>
+          as.integer()
[…]

### 2024-03-03 R/st_manual_join_nb.R

@@ -1,0 +1,136 @@
+#' Manually enforce contiguity between two areas
+#'
+#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
+#' @param x name or number of first area
+#' @param y name or number of second area
+#'
+#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
+#' @export
+#'
+#' @examples
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_manual_join_nb("Gower","St Ives")
+st_manual_join_nb <- function(nb,x,y){
+
+  if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
+    stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
+  } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
+    stop("Error: The dataframe must contain a column called 'nb'")
+  }
+
+  # MAKE INTEGERS: xnum and ynum
+
+  ### case when numeric x and y:
+  if(is.numeric(x) & is.numeric(y)){
+    xnum <- x |> as.integer()
+    ynum <- y |> as.integer()
+  }else{
+
+    ### case when character x and y:
+    # first, make into numeric is given as character
+    # depending on its structure as df, list, matrix...
+    if(is.data.frame(nb)){
+      if(is.list(nb$nb)){
+        xnum <- which(names(nb$nb)==x) |>
+          as.integer()
+        ynum <- which(names(nb$nb)==y) |>
+          as.integer()
+      }
+      if(is.matrix(nb$nb)){
+        xnum <- which(rownames(nb$nb)==x) |>
+          as.integer()
+        ynum <- which(rownames(nb$nb)==y) |>
+          as.integer()
+      }
+    }else{
+      if(is.list(nb)){
+        xnum <- which(names(nb)==x) |>
+          as.integer()
+        ynum <- which(names(nb)==y) |>
+          as.integer()
+      }
+      if(is.matrix(nb)){
+        xnum <- which(rownames(nb)==x) |>
+          as.integer()
+        ynum <- which(rownames(nb)==y) |>
+          as.integer()
+      }
+    }
[…]

### 2024-03-03 R/st_quickmap_nb.R

@@ -1,0 +1,151 @@
+#' Visualise a neighbourhood structure on a map
+#'
+#' @param nbsf an sf dataframe with a neighbourhood column called "nb"
+#' @param linkcol colour of lines connecting neighbours
+#' @param bordercol colour of boundary lines between areas
+#' @param pointcol colour of centroid points if nodes are point
+#' @param fillcol fill of areas
+#' @param linksize linewidth of lines connecting neighbours
+#' @param bordersize linewidth of borders between areas
+#' @param pointsize size of centroid points if nodes are point
+#' @param title plot title
+#' @param subtitle plot subtitle
+#' @param nodes point or numeric
+#' @param numericsize font size if nodes are numeric
+#' @param numericcol font colour if nodes are numeric
+#' @param concavehull default FALSE. whether or not to show concave hulls
+#' @param hullratio fraction convex: 1 returns the convex hulls, 0 maximally concave hulls
+#' @param hullcol colour of concave hull lines
+#' @param hullsize line width of concave hull lines
+#'
+#' @return ggplot of areas and neighbourhood structure
+#' @export
+#'
+#' @examples
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_quickmap_nb()
+st_quickmap_nb <- function(nbsf,
+                           linkcol = "dodgerblue",
+                           bordercol = "gray7",
+                           pointcol="darkred",
+                           fillcol = "gray95",
+                           linksize=0.2,
+                           bordersize=0.1,
+                           pointsize=0.8,
+                           title=NULL,
+                           subtitle=NULL,
+                           nodes="point",
+                           numericsize=5,
+                           numericcol="black",
+                           concavehull=FALSE,
+                           hullratio=0.8,
+                           hullcol="darkgreen",
+                           hullsize=0.5){
+
+  if (!inherits(nbsf,"sf")) {
+    stop("Error: This function requires a simple features dataframe as input")
+  }
+
[…]

### 2024-03-03 R/st_quickmap_preds.R

@@ -1,0 +1,92 @@
+#' Visualilse the predictions of an `mgcv` model
+#'
+#' @param output an augmented sf dataframe produced by st_augment()
+#' @param scale_low fill of lowest extreme of scale
+#' @param scale_mid fill of midpoint of scale
+#' @param scale_high fill of highest extreme of scale
+#' @param scale_midpoint value of midpoint of scale
+#' @param borderwidth linewidth of borders between units
+#' @param bordercol colour of borders between units
+#' @param legendlimits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
+#' @param titlesize font size for title
+#' @param subtitlesize font size for subtitle
+#' @param framefill colour for background fill
+#' @param frameline colour for frame
+#' @param framesize line width of frame
+#'
+#' @return a list of ggplots
+#' @export
+#'
+#' @examples
+#' prepdata <- st_bridges(uk_election,"constituency_name")
+#' mgcv::gam(health_not_good ~
+#'   s(constituency_name, bs='mrf', xt=list(nb=prepdata$nb), k=100),data=prepdata, method="REML") |>
+#' st_augment(uk_election) |>
+#' st_quickmap_preds()
+st_quickmap_preds <- function(output,
+                              scale_low = "firebrick4",
+                              scale_mid = "white",
+                              scale_high = "darkblue",
+                              scale_midpoint = 0,
+                              borderwidth = 0.05,
+                              bordercol = "black",
+                              legendlimits = "individual",
+                              titlesize = 12,
+                              subtitlesize = 10,
+                              framefill = "white",
+                              frameline = "black",
+                              framesize = 1){
+
+  if (!inherits(output,"sf")) {
+    stop("Error: This function requires a simple features dataframe as input")
+  }
+
+  output1 <- output |>
+    dplyr::select(dplyr::starts_with("random.effect"),dplyr::starts_with("mrf.smooth"))
+
+  min_scale <- output1 |>
+    sf::st_drop_geometry() |>
+    min(na.rm = TRUE)
+
+  max_scale <- output1 |>
+    sf::st_drop_geometry() |>
+    max(na.rm = TRUE)
+
+  fillnames <- output1 |>
+    sf::st_drop_geometry() |>
+    names()
+
+  # split column names into title and subtitle
+  # either side of second . in string
+  newsubtitle <- sub("^(.*?\\..*?)\\..*$", "\\1", fillnames)
[…]

### 2024-03-03 R/uk_election.R

@@ -1,0 +1,24 @@
+#' UK election data
+#'
+#' Swing and socio-economic data for England, Scotland & Wales
+#' Census and voting data sourced from parlitools R package
+#' Spatial data sourced from UK government geoportal
+#'
+#' @format ## `uk_election`
+#'    An sf and data.frame object with 632 rows and 9 columns
+#'
+#' \describe{
+#'   \item{degree_educated}{Percentage of constituency population with level 4 qualifications or higher, scaled to mean 0 and standard deviation 1}
+#'   \item{health_not_good}{Percentage of constituency of population reporting health to be fair, bad, or very bad, scaled to mean 0 and standard deviation 1}
+#'   \item{white}{Percentage of constituency of population of exclusively white ethnicity, scaled to mean 0 and standard deviation 1}
+#'   \item{con_swing}{Butler swing to the Conservative Party from the Labour Party from election 2019 to election 2019}
+#'   \item{population}{Constituency population}
+#'   \item{region}{Regions}
+#'   \item{county}{Counties}
+#'   \item{constituency_name}{Westminster parliamentary constituencies, as of 2019}
+#'   \item{geometry}{sfc polygons column}
+#'   ...
+#' }
+#'
+#' @source <https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk>, <https://docs.evanodell.com/parlitools/>
+"uk_election"

### 2024-03-03 data-raw/uk_election.R

@@ -1,0 +1,81 @@
+## code to prepare `uk_election` dataset goes here
+
+library(parlitools)
+library(dplyr)
+library(sf)
+# prepare the data
+# extract and join census and election data from parlitools package
+census_11 <- parlitools::census_11 |>
+  dplyr::select(-constituency_name,-constituency_type,-pano, -region, -country)
+bes_2019 <- parlitools::bes_2019
+
+elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
+
+# download constituency boundaries for 2019 using the st_read() command below
+# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
+uk_map_download <- st_read(
+  "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
+  quiet = TRUE)
+
+
+# only need the boundaries and the IDs for merging with parlitools data
+uk<- uk_map_download |>
+  dplyr::select(pcon19cd,geometry) |>
+  sf::st_transform(crs=27700) |>
+  sf::st_make_valid() # ensure valid line overlaps etc
+
+# create dataframe for filtering-out speaker constituencies below
+speakers <- data.frame(
+  year = c(2017,2019),
+  constituency_name = c("Buckingham","Chorley")
+)
+
+# join constituency polygons to parlitools data
+# filter out Scotland and Northern Ireland
+# filter out the speakers' constituencies
+# calculate the desired explanatory variables
+# make levels into factors for use with `mgcv`
+
+df <- elect_results |>
+  dplyr::left_join(uk, by=c("ons_const_id"="pcon19cd")) |>
+  # filter(!country %in% c("Scotland","Northern Ireland")) |>
+  dplyr::filter(!country %in% c("Northern Ireland")) |>
+  # filter(!constituency_name %in% speakers$constituency_name) |>
+  dplyr::mutate(degree_educated = qual_level_4,
+         health_not_good = health_fair + health_bad + health_very_bad,
+         white = ethnicity_white,
+         con_change = con_19 - con_17, # difference in %
+         lab_change = lab_19 - lab_17,
+         con_swing = (con_change - lab_change)/2, # Butler swing
+         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
[…]

### 2024-03-03 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(sfislands)
+
+test_check("sfislands")

### 2024-03-03 tests/testthat/test-st_bridges.R

@@ -1,0 +1,61 @@
+test_that("multiplication works", {
+  expect_equal(2 * 2, 4)
+})
+
+# list
+nbsf_list <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE)
+nbsf_list_noislands <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE)
+nbsf_list_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE)
+nbsf_list_noislands_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, add_to_dataframe = FALSE)
+
+# matrix
+nbsf_matrix <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, nb_structure = "matrix")
+nbsf_matrix_noislands <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, nb_structure = "matrix")
+nbsf_matrix_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE, nb_structure = "matrix")
+nbsf_matrix_noislands_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, add_to_dataframe = FALSE, nb_structure = "matrix")
+
+test_that("neighbors are of appropriate length for lists", {
+  skip_on_cran()
+  skip_on_ci()
+
+  # list
+  expect_length(nbsf_list, 10)
+  expect_length(nbsf_list$nb, 632)
+  expect_equal(nrow(nbsf_list), 632)
+
+  expect_length(nbsf_list_noislands, 10)
+  expect_length(nbsf_list_noislands$nb, 628)
+  expect_equal(nrow(nbsf_list_noislands), 628)
+
+  expect_length(nbsf_list_nodf, 632)
+  expect_equal(nrow(nbsf_list_nodf), NULL)
+
+  expect_length(nbsf_list_noislands_nodf, 628)
+  expect_equal(nrow(nbsf_list_noislands_nodf), NULL)
+
+  expect_equal(names(nbsf_list$nb),uk_election$constituency_name)
+
+})
+
+test_that("neighbors are of appropriate length for matrices", {
+  skip_on_cran()
+  skip_on_ci()
+
+  # matrix
+  expect_length(nbsf_matrix, 10)
[…]

### 2024-03-03 tests/testthat/test-st_check_islands.R

@@ -1,0 +1,13 @@
+test_that("multiplication works", {
+  expect_equal(2 * 2, 4)
+})
+
+checkdf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE) |>
+  st_check_islands()
+
+test_that("indices match names", {
+  skip_on_cran()
+  skip_on_ci()
+
+  expect_equal(checkdf[,1], uk_election$constituency_name[as.numeric(checkdf[,2])])
+})

### 2024-03-10 R/st_augment.R
punctuation for function title and parameter descriptions

@@ -1,9 +1,9 @@
 #' Augment dataframe with predictions of model
 #'
-#' @param model an `mgcv`, `lme4` or `nlme` model
-#' @param df the sf data frame to be augmented with model predictions
+#' @param model an `mgcv`, `lme4` or `nlme` model.
+#' @param df an `sf` data frame to be augmented with model predictions.
 #'
-#' @return an augmented sf dataframe
+#' @return an augmented `sf` data frame with extra columns showing estimates of random effects from model.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_bridges.R

@@ -1,28 +1,24 @@
 #' Create neighbourhood structure when islands are present
 #'
-#' @param df an sf or sfc object
-#' @param geom_col_name name of column containing names (or unique identifiers) for each row
-#' @param remove_islands TRUE or FALSE (default)
-#' @param link_islands_k an integer, k. Each island should be connected to the k nearest units
-#' @param nb_structure either "list" (default) or "matrix"
-#' @param add_to_dataframe TRUE (default) or FALSE. Whether or not to augment existing df with output as "nb" column
+#' @param df an `sf` or `sfc` object.
+#' @param geom_col_name name of column containing names (or unique identifiers) for each row.
+#' @param remove_islands default `FALSE`. Whether or not to omit islands from contiguity construction.
+#' @param link_islands_k an integer, k. The number of nearest units to which each island should be connected.
+#' @param nb_structure default `"list"`. Can also be `"matrix"`.
+#' @param add_to_dataframe default `TRUE`. Whether or not to augment existing df with contiguity output as `"nb"` column. `FALSE` returns only the contiguity structure.
 #'
-#' @return a named neighbourhood list or matrix, optionally included as a column in an sf dataframe
+#' @return either a named neighbourhood list or matrix, or an `sf` dataframe with list or matrix included as `"nb"` column.
 #' @export
 #'
 #' @examples
 #' st_bridges(uk_election,"constituency_name")
-st_bridges <- function(df, # sf dataframe
+st_bridges <- function(df,
                        geom_col_name,
                        remove_islands = FALSE,
-                       link_islands_k = 1, # link island to k nearest units, 0 removes all islands
+                       link_islands_k = 1,
                        nb_structure = "list",
                        add_to_dataframe = TRUE)
 {
-
-  # if (!inherits(df,"sf") || !inherits(df,"sfc")) {
-  #   stop("Error: df must be an sf or sfc object")
-  # }
 
   if (remove_islands == TRUE){
     # unconnected units

### 2024-03-10 R/st_check_islands.R

@@ -1,8 +1,8 @@
-#' Examine contiguity actions performed on islands
+#' Examine contiguity actions which have been performed on islands by `st_bridges()`
 #'
-#' @param data an sf dataframe with a neighbourhood column called "nb"
+#' @param data an `sf` dataframe with a neighbourhood column called `"nb"` such as the output of `st_bridges()`.
 #'
-#' @return a dataframe reporting non-contiguous connections made by st_bridges()
+#' @return a dataframe reporting non-contiguous connections made by `st_bridges()`.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_manual_cut_nb.R

@@ -1,10 +1,10 @@
 #' Manual remove contiguity between two areas
 #'
-#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
-#' @param x name or number of first area
-#' @param y name or number of second area
+#' @param nb a neighbourhood `"list"` or `"matrix"`, or an `sf` dataframe with a neighbourhood column called `"nb"`.
+#' @param x name or number of first area.
+#' @param y name or number of second area.
 #'
-#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
+#' @return an amended neighbourhood `"list"`, `"matrix"`, or `sf` dataframe with a neighbourhood column called `"nb"`.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_manual_join_nb.R

@@ -1,10 +1,10 @@
 #' Manually enforce contiguity between two areas
 #'
-#' @param nb a neighbourhood list or matrix, or an sf dataframe with a neighbourhood column called "nb"
-#' @param x name or number of first area
-#' @param y name or number of second area
+#' @param nb a neighbourhood `"list"` or `"matrix"`, or an `sf` dataframe with a neighbourhood column called `"nb"`.
+#' @param x name or number of first area.
+#' @param y name or number of second area.
 #'
-#' @return an amended neighbourhood list, matrix, or sf dataframe with a neighbourhood column called "nb"
+#' @return an amended neighbourhood `"list"`, `"matrix"`, or `sf` dataframe with a neighbourhood column called `"nb"`.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_quickmap_nb.R

@@ -1,24 +1,24 @@
 #' Visualise a neighbourhood structure on a map
 #'
-#' @param nbsf an sf dataframe with a neighbourhood column called "nb"
-#' @param linkcol colour of lines connecting neighbours
-#' @param bordercol colour of boundary lines between areas
-#' @param pointcol colour of centroid points if nodes are point
-#' @param fillcol fill of areas
-#' @param linksize linewidth of lines connecting neighbours
-#' @param bordersize linewidth of borders between areas
-#' @param pointsize size of centroid points if nodes are point
-#' @param title plot title
-#' @param subtitle plot subtitle
-#' @param nodes point or numeric
-#' @param numericsize font size if nodes are numeric
-#' @param numericcol font colour if nodes are numeric
-#' @param concavehull default FALSE. whether or not to show concave hulls
-#' @param hullratio fraction convex: 1 returns the convex hulls, 0 maximally concave hulls
-#' @param hullcol colour of concave hull lines
-#' @param hullsize line width of concave hull lines
+#' @param nbsf an `sf` dataframe with a neighbourhood column called "nb", such as the output of `st_bridges()`
+#' @param linkcol colour of lines connecting neighbours.
+#' @param bordercol colour of boundary lines between areas.
+#' @param pointcol colour of centroid points if nodes are `"point"`.
+#' @param fillcol fill of areas.
+#' @param linksize linewidth of lines connecting neighbours.
+#' @param bordersize linewidth of borders between areas.
+#' @param pointsize size of centroid points if nodes are `"point"`.
+#' @param title plot title.
+#' @param subtitle plot subtitle.
+#' @param nodes default `"point"`. Can also be `"numeric"`.
+#' @param numericsize font size if nodes are `"numeric"`.
+#' @param numericcol font colour if nodes are `"numeric"`.
+#' @param concavehull default `FALSE`. Whether or not to show concave hulls.
+#' @param hullratio value between 0 and 1. 1 returns the convex hulls, 0 maximally concave hulls.
+#' @param hullcol colour of concave hull lines.
+#' @param hullsize line width of concave hull lines.
 #'
-#' @return ggplot of areas and neighbourhood structure
+#' @return `ggplot` of areas and neighbourhood structure.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_quickmap_preds.R

@@ -1,20 +1,20 @@
-#' Visualilse the predictions of an `mgcv` model
+#' Visualise the predictions generated by the `st_augment()` function
 #'
-#' @param output an augmented sf dataframe produced by st_augment()
-#' @param scale_low fill of lowest extreme of scale
-#' @param scale_mid fill of midpoint of scale
-#' @param scale_high fill of highest extreme of scale
-#' @param scale_midpoint value of midpoint of scale
-#' @param borderwidth linewidth of borders between units
-#' @param bordercol colour of borders between units
-#' @param legendlimits "individual" (default) each plot scaled to individual min-max. "minmax" means all plot have a common min-max according to the global min-max
-#' @param titlesize font size for title
-#' @param subtitlesize font size for subtitle
-#' @param framefill colour for background fill
-#' @param frameline colour for frame
-#' @param framesize line width of frame
+#' @param output an augmented `sf` dataframe produced by `st_augment()`.
+#' @param scale_low fill of lowest extreme of scale.
+#' @param scale_mid fill of midpoint of scale.
+#' @param scale_high fill of highest extreme of scale.
+#' @param scale_midpoint value of midpoint of scale.
+#' @param borderwidth linewidth of borders between units.
+#' @param bordercol colour of borders between units.
+#' @param legendlimits default `"individual"`. legend of each plot scaled within its own limits. `"minmax"` means all plot have common legend limits according to the global min-max.
+#' @param titlesize font size for title.
+#' @param subtitlesize font size for subtitle.
+#' @param framefill colour for background fill.
+#' @param frameline colour for frame.
+#' @param framesize line width of frame.
 #'
-#' @return a list of ggplots
+#' @return a `"list"` of ggplots.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_augment.R
further changes from checklist

@@ -3,7 +3,7 @@
 #' @param model an `mgcv`, `lme4` or `nlme` model.
 #' @param df an `sf` data frame to be augmented with model predictions.
 #'
-#' @return an augmented `sf` data frame with extra columns showing estimates of random effects from model.
+#' @return An augmented `sf` data frame with extra columns showing estimates of random effects from model.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_bridges.R

@@ -1,13 +1,13 @@
-#' Create neighbourhood structure when islands are present
+#' Creates first-order queen contiguity neighbourhood structure with additional connections when islands are present, ensuring that there are no unconnected units
 #'
 #' @param df an `sf` or `sfc` object.
-#' @param geom_col_name name of column containing names (or unique identifiers) for each row.
+#' @param geom_col_name name of a column from `df` containing names (or unique identifiers) for each row.
 #' @param remove_islands default `FALSE`. Whether or not to omit islands from contiguity construction.
 #' @param link_islands_k an integer, k. The number of nearest units to which each island should be connected.
-#' @param nb_structure default `"list"`. Can also be `"matrix"`.
+#' @param nb_structure default `"list"`. Can also be `"matrix"`. The format in which to return the named contiguity structure.
 #' @param add_to_dataframe default `TRUE`. Whether or not to augment existing df with contiguity output as `"nb"` column. `FALSE` returns only the contiguity structure.
 #'
-#' @return either a named neighbourhood list or matrix, or an `sf` dataframe with list or matrix included as `"nb"` column.
+#' @return Either a named neighbourhood list or matrix, or an `sf` dataframe with list or matrix included as `"nb"` column.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_check_islands.R

@@ -2,7 +2,7 @@
 #'
 #' @param data an `sf` dataframe with a neighbourhood column called `"nb"` such as the output of `st_bridges()`.
 #'
-#' @return a dataframe reporting non-contiguous connections made by `st_bridges()`.
+#' @return A dataframe reporting non-contiguous connections made by `st_bridges()`.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_manual_join_nb.R

@@ -4,7 +4,7 @@
 #' @param x name or number of first area.
 #' @param y name or number of second area.
 #'
-#' @return an amended neighbourhood `"list"`, `"matrix"`, or `sf` dataframe with a neighbourhood column called `"nb"`.
+#' @return An amended neighbourhood `"list"`, `"matrix"`, or `sf` dataframe with a neighbourhood column called `"nb"`.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_quickmap_nb.R

@@ -18,7 +18,7 @@
 #' @param hullcol colour of concave hull lines.
 #' @param hullsize line width of concave hull lines.
 #'
-#' @return `ggplot` of areas and neighbourhood structure.
+#' @return A `ggplot` showing areas and neighbourhood structure.
 #' @export
 #'
 #' @examples

### 2024-03-10 R/st_quickmap_preds.R

@@ -14,7 +14,7 @@
 #' @param frameline colour for frame.
 #' @param framesize line width of frame.
 #'
-#' @return a `"list"` of ggplots.
+#' @return A list of ggplots.
 #' @export
 #'
 #' @examples

### 2024-03-14 README.Rmd
this?
further changes from checklist
improved readme

--- 2024-02-07 README.Rmd
@@
 
+<img src="man/figures/logo.png" align="right" height="240"/>
+
 # sfislands
@@
 
--   These do not have to be "literal" islands but any situation where discontiguous geographical units are present.
+-   These do not have to be *islands* by a traditional *land and water* geographic definition, but any situation where discontiguous geographical units are present.
 
@@
 
+Install the released version from CRAN
+
+``` r
+install.packages("sfislands")
+```
+
 You can install the development version of `sfislands` from [GitHub](https://github.com/horankev/sfislands) with:
@@
 1.
-The initial setting up neighbourhood structures can be frustrating for people who are eager to get started with fitting spatial models.
+The initial setting-up of neighbourhood structures can be frustrating for people who are eager to get started with fitting spatial models.
 This is especially so when the presence of discontiguities within a geographical dataset means that, even having set up a neighbours list, the model will still not run without further awkward data manipulations.
@@
 
-3.  If there are some neighbours assigned which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.
+3.
+If some units have been assigned as neighbours inappropriately, or if we wish to add additional connections, there are functions to allow this to be done in a straightforward and openly reportable way.
 
@@
 Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed.
-`sfdep` contains functionality to perform such test, and the output from `sfislands` can be used in its functions.
+`sfdep` contains functionality to perform such tests, and the output from `sfislands` can be used in its functions.
 
@@
 
-<img src="man/figures/logo.png" align="right" height="240"/>
-
 ## Functions overview
 
-The following is a framework within which the `sfislands` functions could be used.
+The following is a brief explanations of the functions contained in the `sfislands` package and a workflow within which they could be used.
 
@@
 
-*Use the output of **st_bridges()** as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.*
+Use the output of `st_bridges()` as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.
 
@@
 ```
-
-## Examples
-
-In the context of the constituencies of England, Scotland and Wales, `st_bridges` will define a neighbourhood structure which includes the islands.
-`st_quickmap_nb` will show a representation of this on a map:
-
 ```{r, echo=FALSE}
@@
 
-```{r, fig.width=12, fig.height=10}
-nbsf <- st_bridges(df = uk_election,
+## Workflow
+
+The dataset `uk_elections`, which is provided with the package, is used for the following demonstrations.
+It is an `sf` dataset with geometries for each of the 632 Westminster parliamentary constituencies of England, Scotland and Wales in the 2019 general election.
+For each constituency, it contains socio-economic data from the 2011 census.
+It also shows the *Butler swing* between the 2017 and 2019 elections, which is a measure of the change in voter allegiance in favour of the Conservative Party from the Labour Party.
+
+### Pre-functions: Creating desired adjacency structure
+
+To demonstrate the `st_bridges()` function for preparing an adjacency matrix, we focus on Scotland.
+It is the nation with the highest number of islands.
+The function, with default arguments, returns a dataframe with an additional column named `nb`.
+This column is a named neighbourhood list of class `nb`.
+The default neighbourhood structure which it computes is:
+
+1.    for constituencies which are **not islands**, first-order queen contiguity. A constituency is considered to be a neighbour of all of those with which it shares at least a vortex.
+
+2.    for **islands**, contiguities are assigned to the two nearest constituencies to each island. This number can be altered using the `link_islands_k` argument.
+
+```{r}
+nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
                    geom_col_name = "constituency_name",
                    link_islands_k = 2)
+```
+
+This new dataframe has the following form (with a column named `nb` containing a neighbours list):
+
+```{r}
+head(nbsf)
+```
+We can inspect (and openly report) what additional contiguities have been added to the standard queen contiguity structure using the function `st_check_islands()`.
+
+```{r}
+st_check_islands(nbsf)
+```
+
+The overall contiguity structure can be visualised using the function `st_quickmap_nb()`, whose output is shown below on the left.
+
+```{r, fig.width=12, fig.height=10}
+ggarrange(
+  st_quickmap_nb(nbsf),
+
   st_quickmap_nb(nbsf,
-                    pointsize=0.05, 
-                    title = "st_bridges() contiguities",
-                    subtitle = "islands linked to k=2 nearest constituencies")
+                 concavehull = TRUE,
+                 nodes = "numeric",
+                 numericsize = 3.5),
+
+  ncol = 2
+)
 
@@
 
-Neighbours can be manually added or removed using `st_manual_join_nb()` and `st_manual_cut_nb()`.
+However, it is often the case that each *island* unit (in this case, constituency) is in fact a multipolygon of islands.
+To make this clearer, the argument `concavehull = TRUE` can be chosen.
+These green boundaries (shown above on the right) around each constituency are for reference purpose only and do not influence the assignment of contiguities.
+They show which group of *islands* are considered as one *constituency of islands*.
+Furthermore, the argument `nodes = "numeric"` represents the centroids of each unit with its index number instead of a geometric point.
+This can be useful for editing the adjacencies, particularly if the researcher is not familiar with the names of the individual spatial units.
+
+#### Editing contiguities
+
+A researcher can now use domain knowledge or a particular hypothesis to edit these contiguities.
+For instance, we may not be happy with the connection between 47 (Na h-Eileanan An Iar) and 51 (Orkney and Shetland) and wish to remove it.
+We might instead feel that it is important for 47 and 5 to be considered neighbours.
+This can be easily achieved, as shown below, using the `st_manual_cut_nb()` and `st_manual_join_nb()` functions.
+The spatial units whose relationships we wish to change can be referenced by name or by index number.
 
 ```{r}
-st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
+nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
                    geom_col_name = "constituency_name",
                    link_islands_k = 2) |> 
-  st_manual_join_nb("Gower","St Ives") |> 
-  st_quickmap_nb()
+  st_manual_cut_nb("Na h-Eileanan An Iar", "Orkney and Shetland") |> 
+  st_manual_join_nb(47, 5)
 ```
 
-## Modelling & post-functions
+The results of these changes as visualised in map form are shown below:
 
-Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.
+```{r, fig.width=12, fig.height=10}
+ggarrange(
+  st_quickmap_nb(nbsf,
+                 concavehull = TRUE,
+                 nodes = "numeric",
+                 numericsize = 3.5),
 
-For example, we can use the `mgcv` package to generate a Markov Random Field ICAR smooth of poor health across the study area.
-This is done very quickly by using `st_bridges()` to prepare the data, putting that inside the `mgcv` GAM formulation, and then piping into the `st_augment()` and `st_quickmap_preds()` functions.
+  st_quickmap_nb(nbsf),
 
-```{r, fig.width=12, fig.height=6}
+  ncol = 2
+)
 
-prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
+```
 
-model <- gam(con_swing ~ 
+The above process can be iterated until we are happy that the structure conforms with our hypothesis of connectivity within the study area.
+
+### Modelling & post-functions
+
+Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as a column named `nb`, the workflow to generate and visualise predictions from a model (in this case using `mgcv`) is very straightforward and consists of only 3-4 lines of code.
+
+For example, we can use the `mgcv` package to generate quite a complicated model with various random intercepts and slopes at region and county level, and additionally, different Markov Random Field ICAR smooths at the constituency level.
+This is done very quickly by
+
+* using `st_bridges()` to prepare the data, 
+* putting that inside the `mgcv::gam()` formulation, and then 
+* piping into the `st_augment()` function.
+
+This gives the following output, with all of the spatially varying predictions and their standard errors being added as extra columns before the final geometry column:
+
+```{r}
+prep_data <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
+
+model1 <- gam(con_swing ~ 
                s(region, bs="re") + # region level random intercept
@@
                s(constituency_name, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
+                 xt=list(nb=prep_data$nb),k=10) + # ICAR constituency ICAR varying coefficients
                s(constituency_name, by=white, bs='mrf', 
-                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
-             data=prep_data3, method="REML") |> 
-  st_augment(prep_data3) |> # pipe into function to get estimates
-  st_quickmap_preds() # pipe into this for visualisation
+                 xt=list(nb=prep_data$nb),k=10), # ICAR constituency ICAR varying coefficients
+             data=prep_data, method="REML") |>  # pipe into function to get estimates
+  st_augment(prep_data)
 
-ggarrange(plotlist = model, legend = "none", nrow=1)
+head(model1)
+
 ```
 
-Further information and worked examples are available in the vignettes which can be found [here](https://horankev.github.io/sfislands/).
+The predictions can be quickly visualised using `st_quickmap_preds()`.
+As this function generates a list of `ggplots`, they can be conveniently visualised using the `ggarrange()` function from the `ggpubr` package.
+
+```{r, fig.width=12, fig.height=6}
+plots <- st_quickmap_preds(model1)
+ggarrange(plotlist = plots, legend = "none", nrow=1)
+```
+
+To emphasise the simplicity of the process, the a modified version of the following sequence of commands will generate a series of plots of all of the spatially varying elements of a model defined in `mgcv`:
+
+```{r, eval=FALSE}
+prep_data <- st_bridges(...)
+
+gam(..., data = prep_data) |> 
+  st_augment() |> 
+  st_quickmap_preds()
+
+```
+
+It is also possible to conduct the same kind of augmenting and plotting of models generated using the `lme4` and `nlme` packages.
+
+Further information and vignettes containing more detailed examples are available [here](https://horankev.github.io/sfislands/).

### 2024-03-14 R/st_bridges.R

@@ -1,4 +1,4 @@
-#' Creates first-order queen contiguity neighbourhood structure with additional connections when islands are present, ensuring that there are no unconnected units
+#' Create first-order queen contiguity neighbourhood structure with additional connections when islands are present, ensuring that there are no unconnected units
 #'
 #' @param df an `sf` or `sfc` object.
 #' @param geom_col_name name of a column from `df` containing names (or unique identifiers) for each row.
@@ -95,6 +95,8 @@ st_bridges <- function(df,
       tempdf <- df |>
         dplyr::mutate(nb = cont)
       tempdf[[geom_col_name]] <- factor(tempdf[[geom_col_name]])
+      tempdf <- tempdf |>
+        dplyr::select(everything(),nb,geometry)
       return(tempdf)
     }
     if(nb_structure =="matrix"){
@@ -103,6 +105,8 @@ st_bridges <- function(df,
       tempdf <- df |>
         dplyr::mutate(nb = cont2)
       tempdf[[geom_col_name]] <- factor(tempdf[[geom_col_name]])
+      tempdf <- tempdf |>
+        dplyr::select(everything(),nb,geometry)
       return(tempdf)
     }
   }

### 2024-03-15 R/st_augment.R
ordering of augment function mgcv

@@ -88,7 +88,10 @@ st_augment.gam <- function(model,df){
   output2 <- cbind(output,df) |>
     as.data.frame() |>
     dplyr::select(-dplyr::matches("fit\\.")) |>
-    dplyr::select(-(1:ncol(output)), dplyr::everything()) |>
+    dplyr::select(!dplyr::contains("mrf."),!dplyr::contains("random.effect."),
+                  dplyr::everything(),
+                  dplyr::contains("random.effect."),dplyr::contains("mrf."),
+                  geometry) |>
     sf::st_as_sf()
 
   return(output2)

### 2024-03-15 R/st_augment.R
col sequence for mgcv st_augment

@@ -88,9 +88,11 @@ st_augment.gam <- function(model,df){
   output2 <- cbind(output,df) |>
     as.data.frame() |>
     dplyr::select(-dplyr::matches("fit\\.")) |>
-    dplyr::select(!dplyr::contains("mrf."),!dplyr::contains("random.effect."),
+    # dplyr::select(!dplyr::contains("mrf.") & !dplyr::contains("random.effect."),
+    #               dplyr::contains("random.effect."),dplyr::contains("mrf."),
+    #               geometry) |>
+    dplyr::select(names(df),
                   dplyr::everything(),
-                  dplyr::contains("random.effect."),dplyr::contains("mrf."),
                   geometry) |>
     sf::st_as_sf()
 

### 2024-04-16 README.Rmd
hex position on readme
with adjust col seq for st_augment
minor alteration
add arxiv link to readme

@@
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
 
-<img src="man/figures/logo.png" align="right" height="240"/>
+# sfislands <img src="man/figures/logo.png" align="right" height="240"/>
 
-# sfislands
+<!-- badges: start -->
+<!-- badges: end -->
 
@@
 
+A pre-print of a paper discussing the package is available [here](https://arxiv.org/abs/2404.09863).
+
 ## Installation
 
-Install the released version from CRAN
+Install the released version from CRAN:
 

### 2025-02-28 R/st_bridges.R
further refinements

@@ -6,6 +6,8 @@
 #' @param link_islands_k an integer, k. The number of nearest units to which each island should be connected.
 #' @param nb_structure default `"list"`. Can also be `"matrix"`. The format in which to return the named contiguity structure.
 #' @param add_to_dataframe default `TRUE`. Whether or not to augment existing df with contiguity output as `"nb"` column. `FALSE` returns only the contiguity structure.
+#' @param threshold default 1.001. factor by which to change the size of buffer automatically generated around islands to account for imprecisions which may arise from sf::st_buffer() functionality.
+#' @param geom_col_name name of a column from `df` containing names (or unique identifiers) for each row. This argument is now deprecated and the new "row_identifier" argument replaces it with the same functionality.
 #'
 #' @return Either a named neighbourhood list or matrix, or an `sf` dataframe with list or matrix included as `"nb"` column.
 #' @export
@@ -18,6 +20,7 @@ st_bridges <- function(df,
                        link_islands_k = 1,
                        nb_structure = "list",
                        add_to_dataframe = TRUE,
+                       threshold = 1.001,
                        geom_col_name = lifecycle::deprecated())
 {
 
@@ -58,7 +61,7 @@ st_bridges <- function(df,
     {
       distances <- sf::st_distance(df$geometry[unconnected[i]], sf::st_geometry(df$geometry)) |>
         as.numeric() |> sort()
-      distdf[i,2] <- distances[link_islands_k+1] * 1.001
+      distdf[i,2] <- distances[link_islands_k+1] * threshold
       distdf[i,1] <- unconnected[i]
     }
     bufs <- rep(0,nrow(df))
@@ -71,6 +74,13 @@ st_bridges <- function(df,
       sf::st_buffer(dist=bufs) |>
       sf::st_intersects() |>
       purrr::imap(~setdiff(.x,.y))
+
+    # attempt to use st_is_within_distance instead of buffer and intersects
+    # cont <- purrr::map2(seq_len(nrow(df)), bufs, function(i, buf) {
+    #   # Apply st_is_within_distance for the ith row and buffer distance buf
+    #   sf::st_is_within_distance(df[i, ], df, dist = buf)
+    # }) |>
+    #   purrr::imap(~setdiff(.x, .y))  # Remove self-neighbor (if any)
 
     names(cont) <- df |> dplyr::pull({{ row_identifier }})
 

### 2025-02-28 R/st_check_islands.R

@@ -30,7 +30,14 @@ st_check_islands <- function(data){
   # unconnected units
   cont <-  data |>
     sf::st_intersects()
-  still_unconnected <- lengths(lapply(cont, function(x) x))
+  still_unconnected <- lengths(cont)
+
+  # check if original structure had any islands
+  if(all(still_unconnected > 1)){
+    message("No disconnected units were found in original data")
+    return(0)
+  }
+
   unconnected <- which(still_unconnected == 1)
 
 

### 2025-02-28 R/st_force_join_nb.R

@@ -10,12 +10,12 @@
 #'
 #' @examples
 #' # For individual x and y
-#' st_fj(uk_election, x = "Gower", y = "Cardiff")
+#' st_force_join_nb(uk_election, x = "Gower", y = "Cardiff")
 #'
 #' # For multiple x and y pairs
 #' xy_df <- data.frame(x = c("Gower", "St Ives"), y = c("Cardiff", "Penzance"))
-#' st_fj(uk_election, xy_df = xy_df)
-st_fj <- function(nb, x = NULL, y = NULL, xy_df = NULL){
+#' st_force_join_nb(uk_election, xy_df = xy_df)
+st_force_join_nb <- function(nb, x = NULL, y = NULL, xy_df = NULL){
 
   # Check if nb is a valid neighbourhood list, matrix, or dataframe containing 'nb' column
   if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {

### 2025-02-28 R/st_quickmap_preds.R

@@ -78,7 +78,8 @@ st_quickmap_preds <- function(output,
                                                 else(NULL)) +
       ggplot2::labs(title=newtitle[i],
                     subtitle=newsubtitle[i]) +
-      ggplot2::coord_sf(datum=NA) +
+      ggplot2::coord_sf(datum=NA,
+                        default = TRUE) +#to deal with unwanted 'coordinate system already present' message
       ggplot2::theme(plot.title = ggplot2::element_text(size=titlesize),
                      plot.subtitle = ggplot2::element_text(size=subtitlesize),
                      legend.title =  ggplot2::element_blank(),

### 2025-02-28 README.Rmd
update readme files

@@
 | **st_check_islands()**  | *Check assignment of island contiguities in a dataframe.*                                                                                  |
-| **st_manual_join_nb()** | *Make manual changes to any connections.*                                                                                                  |
-| **st_manual_cut_nb()**  | *Make manual changes to any connections.*                                                                                                  |
+| **st_force_join_nb()** | *Enforce changes to any connections.*                                                                                                  |
+| **st_force_cut_nb()**  | *Enforce changes to any connections.*                                                                                                  |
 
@@
 ```
+
 ```{r, echo=FALSE}
@@
 nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
-                   geom_col_name = "constituency_name",
+                   row_identifier = "constituency_name",
                    link_islands_k = 2)
@@
 nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
-                   geom_col_name = "constituency_name",
+                   row_identifier = "constituency_name",
                    link_islands_k = 2) |> 
-  st_manual_cut_nb("Na h-Eileanan An Iar", "Orkney and Shetland") |> 
-  st_manual_join_nb(47, 5)
+  st_force_cut_nb("Na h-Eileanan An Iar", "Orkney and Shetland") |> 
+  st_force_join_nb(47, 5)
 ```

### 2025-02-28 tests/testthat/test-st_bridges.R
update tests

@@ -3,16 +3,16 @@ test_that("multiplication works", {
 })
 
 # list
-nbsf_list <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE)
-nbsf_list_noislands <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE)
-nbsf_list_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE)
-nbsf_list_noislands_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, add_to_dataframe = FALSE)
+nbsf_list <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = FALSE)
+nbsf_list_noislands <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = TRUE)
+nbsf_list_nodf <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE)
+nbsf_list_noislands_nodf <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = TRUE, add_to_dataframe = FALSE)
 
 # matrix
-nbsf_matrix <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, nb_structure = "matrix")
-nbsf_matrix_noislands <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, nb_structure = "matrix")
-nbsf_matrix_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE, nb_structure = "matrix")
-nbsf_matrix_noislands_nodf <- st_bridges(uk_election, "constituency_name", remove_islands = TRUE, add_to_dataframe = FALSE, nb_structure = "matrix")
+nbsf_matrix <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = FALSE, nb_structure = "matrix")
+nbsf_matrix_noislands <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = TRUE, nb_structure = "matrix")
+nbsf_matrix_nodf <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = FALSE, add_to_dataframe = FALSE, nb_structure = "matrix")
[…]

### 2025-02-28 tests/testthat/test-st_check_islands.R

@@ -2,7 +2,7 @@ test_that("multiplication works", {
   expect_equal(2 * 2, 4)
 })
 
-checkdf <- st_bridges(uk_election, "constituency_name", remove_islands = FALSE) |>
+checkdf <- st_bridges(uk_election, row_identifier = "constituency_name", remove_islands = FALSE) |>
   st_check_islands()
 
 test_that("indices match names", {

### 2025-02-28 R/st_quickmap_nb.R
add ::

@@ -86,7 +86,7 @@ st_quickmap_nb <- function(nbsf,
   # Add numeric labels if nodes = "numeric"
   if (nodes == "numeric") {
     nbsf$id <- seq_len(nrow(nbsf))
-    st_agr(nbsf) <- "constant" #explicitly make attribute constant despite the following geometry operations to avoid warnings
+    sf::st_agr(nbsf) <- "constant" #explicitly make attribute constant despite the following geometry operations to avoid warnings
     plot <- plot + ggplot2::geom_sf_text(data = sf::st_centroid(nbsf),
                                          ggplot2::aes(label = id),
                                          size = numericsize, colour = numericcol, fontface = "bold")

### 2025-02-28 R/st_force_join_nb.R
fix examples cut and join

@@ -10,11 +10,16 @@
 #'
 #' @examples
 #' # For individual x and y
-#' st_force_join_nb(uk_election, x = "Gower", y = "Cardiff")
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_force_join_nb(x = "Gower", y = "Bridgend")
 #'
 #' # For multiple x and y pairs
-#' xy_df <- data.frame(x = c("Gower", "St Ives"), y = c("Cardiff", "Penzance"))
-#' st_force_join_nb(uk_election, xy_df = xy_df)
+#' xy_df <- tibble(
+#' ~x, ~y,
+#' "Gower", "Bridgend",
+#' "LLanelli", "Vale Of Glamorgan")
+#' st_bridges(uk_election,"constituency_name") |>
+#' st_force_join_nb(xy_df = xy_df)
 st_force_join_nb <- function(nb, x = NULL, y = NULL, xy_df = NULL){
 
   # Check if nb is a valid neighbourhood list, matrix, or dataframe containing 'nb' column

### 2025-03-01 R/st_bridges.R
catch duplicate row identifiers

@@ -29,7 +29,7 @@ st_bridges <- function(df,
     row_identifier <- geom_col_name
   }
 
-  if (anyDuplicated(df$row_identifier) != 0) {
+  if (length(unique(df$row_identifier)) != nrow(df)) {
     stop(sprintf("Error: duplicate row identifiers present"), call. = FALSE)
   }
 

### 2025-03-01 R/st_bridges.R
catch duplicate row identifiers

@@ -30,7 +30,7 @@ st_bridges <- function(df,
   }
 
   if (length(unique(df$row_identifier)) != nrow(df)) {
-    stop(sprintf("Error: duplicate row identifiers present"), call. = FALSE)
+    stop(sprintf("duplicate row identifiers present"), call. = FALSE)
   }
 
   if (remove_islands == TRUE){

### 2025-03-01 README.Rmd
fixed duplicate row ident issue

@@
 ```{r}
-nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
+scotland_df <- uk_election |> filter(country == "Scotland")
+nbsf <- st_bridges(df = scotland_df,
                    row_identifier = "constituency_name",

### 2025-03-01 R/st_bridges.R

@@ -29,8 +29,8 @@ st_bridges <- function(df,
     row_identifier <- geom_col_name
   }
 
-  if (length(unique(df$row_identifier)) != nrow(df)) {
-    stop(sprintf("duplicate row identifiers present"), call. = FALSE)
+  if (length(unique(df |> dplyr::pull({{ row_identifier }}))) != nrow(df)) {
+    stop("Duplicate row identifiers present", call. = FALSE)
   }
 
   if (remove_islands == TRUE){

### 2025-03-01 R/st_force_cut_nb.R
cut and join improve checks

@@ -19,136 +19,132 @@
 #' x = c("Llanelli", "Swansea West"),
 #' y = c("Bridgend", "Vale Of Glamorgan")))
 st_force_cut_nb <- function(nb, x = NULL, y = NULL, xy_df = NULL) {
-
-  # Check if nb is a valid neighbourhood list, matrix, or dataframe containing 'nb' column
+  # Check if nb is valid
   if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
     stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
   } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
     stop("Error: The dataframe must contain a column called 'nb'")
   }
 
-  # If xy_df is provided, process multiple x, y pairs from the dataframe
+  # Determine number of areas (length of neighbourhood list/matrix)
+  nb_length <- if (is.data.frame(nb)) {
+    if (is.list(nb$nb)) {
+      length(nb$nb)
+    } else if (is.matrix(nb$nb)) {
+      nrow(nb$nb)
+    } else {
+      stop("Error: Invalid neighbourhood structure in dataframe.")
+    }
+  } else if (is.list(nb)) {
+    length(nb)
+  } else if (is.matrix(nb)) {
+    nrow(nb)
+  } else {
+    stop("Error: Unrecognised neighbourhood format.")
+  }
+
+  # Helper: check validity of x, y pair
+  check_pair_validity <- function(x, y) {
+    if (is.character(x) && is.character(y)) {
+      valid_names <- if (is.data.frame(nb)) {
+        if (is.list(nb$nb)) {
+          names(nb$nb)
+        } else if (is.matrix(nb$nb)) {
+          rownames(nb$nb)
+        }
+      } else if (is.list(nb)) {
+        names(nb)
+      } else if (is.matrix(nb)) {
+        rownames(nb)
+      }
+      if (!(x %in% valid_names)) stop(paste0("Error: x = '", x, "' not found in neighbourhood."))
+      if (!(y %in% valid_names)) stop(paste0("Error: y = '", y, "' not found in neighbourhood."))
+    } else if (is.numeric(x) && is.numeric(y)) {
[…]

### 2025-03-01 R/st_force_join_nb.R

@@ -18,128 +18,6 @@
 #' st_force_join_nb(xy_df = data.frame(
 #' x = c("Gower", "Llanelli"),
 #' y = c("Bridgend", "Vale Of Glamorgan")))
-# st_force_join_nb <- function(nb, x = NULL, y = NULL, xy_df = NULL){
-#
-#   # Check if nb is a valid neighbourhood list, matrix, or dataframe containing 'nb' column
-#   if (!(is.data.frame(nb) || is.list(nb$nb) || is.matrix(nb$nb) || is.list(nb) || is.matrix(nb))) {
-#     stop("Error: The 'nb' argument must be a neighbours list, a neighbours matrix, or a dataframe containing a neighbours list or matrix named 'nb'")
-#   } else if (is.data.frame(nb) && !("nb" %in% colnames(nb))) {
-#     stop("Error: The dataframe must contain a column called 'nb'")
-#   }
-#
-#   # If xy_df is provided, process multiple x, y pairs from the dataframe
-#   if (!is.null(xy_df)) {
-#     # Ensure that xy_df has x and y columns
-#     if (!all(c("x", "y") %in% colnames(xy_df))) {
-#       stop("Error: The dataframe must contain columns 'x' and 'y'")
-#     }
-#
-#     # Loop over the rows of xy_df and perform the join for each pair
-#     for (i in 1:nrow(xy_df)) {
-#       x <- xy_df$x[i]
-#       y <- xy_df$y[i]
-#
-#       # Process the current pair
-#       nb <- process_pair(nb, x, y)
-#     }
-#
-#     return(nb)
-#   }
-#
-#   # If no xy_df, proceed with single x and y processing
-#   if (is.null(x) || is.null(y)) {
-#     stop("Error: Either provide 'x' and 'y' or 'xy_df'.")
-#   }
-#
-#   # Process the single pair
-#   nb <- process_pair(nb, x, y)
-#
-#   return(nb)
-# }
-#
-# # Helper function to process a single pair of x and y
-# process_pair <- function(nb, x, y) {
-#
-#   # MAKE INTEGERS: xnum and ynum
-#   if(is.numeric(x) & is.numeric(y)){
-#     xnum <- as.integer(x)
-#     ynum <- as.integer(y)
-#   }else{
-#     if(is.data.frame(nb)){
-#       if(is.list(nb$nb)){
[…]

### 2025-03-01 R/st_quickmap_nb.R
CRAN required fixes

@@ -90,7 +90,7 @@ st_quickmap_nb <- function(nbsf,
     centroids <- sf::st_centroid(nbsf)
     centroids$id <- nbsf$id
     plot <- plot + ggplot2::geom_sf_text(data = centroids,
-                                         ggplot2::aes(label = id),
+                                         ggplot2::aes(label = nbsf$id),
                                          size = numericsize, colour = numericcol, fontface = "bold")
   } else {
     plot <- plot + ggplot2::geom_sf(data = endpoints_coords, size = pointsize, colour = pointcol)
